### 2016-06-06 paper.tex
initial

\documentclass[nocopyrightspace]{sigplanconf}

\usepackage{amsmath}
\usepackage{color}
\usepackage[svgnames,x11names]{xcolor}
\usepackage[hyphens]{url}
\usepackage[citecolor=DarkGreen,linkcolor=DarkGreen,urlcolor=DarkGreen,colorlinks]{hyperref}
\usepackage{epsfig}

\newcommand{\TODO}[1]{\textbf{[TODO: #1]}}

\begin{document}
\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\title{Conex - establishing trust into repositories}

\authorinfo{Hannes Mehnert\thanks{Funded by EPSRC Programme Grant EP/K008528/1}}{University of Cambridge}{}
\authorinfo{Louis Gesbert}{OCamlPro}{}

\date{}

\maketitle

\begin{abstract}
  Software update systems, such as opam, should verify that packages and metadata thereof are authenticated.
  Opam does not include any mechanism to verify authenticity at the moment.
  We propose \emph{conex}, which uses end-to-end digital signatures from the package author to the user.
  Conex is implemented in OCaml and can be integrated into opam to build a security story where even the central repository server does not need to be trusted.
  Authors and users will not need to change their workflow, whereas janitors (repository maintainers) will have to adjust their workflow.
\end{abstract}

\section{Introduction}
Software update systems, such as opam, need to discover, download, and install packages.
Crucial to all steps is that the received data is authenticated before being used to prevent attacks injecting backdoors.
A software update system relies on a central server which provides a repository (source code and metadata such as dependencies and build recipe) of all packages.
When a new package is released, the repository is updated.
The central server is an online services and a single point of failures, and recently these servers been attacked \cite{adobe,apache2,apache1,opera,kernel,cabal,kernelorg,debian,freebsd,php3,savannah,maven,savannah2,npm,php,phpmyadmin,php2,github,ocamlforge,redhatssh,gentoo,rubygems,fedora,wordpress} to distribute malware.

While so far there is no known breach of opam installations (but other package managers have been compromised, see \cite{cabal,maven,npm,rubygems}), there is not much security in place:
In opam 1.2.2 none of the steps (discover, download, install) includes verification.
The discover and download steps use https, but pass \texttt{--no-check-certificates} to the download utility \texttt{wget}, which disables certificate verification.
Any man-in-the-middle between client and repository can arbitrarily modify updates or tarballs.
The build and install steps may do any modifications to the file system, not limited to the build and target directory of the package.
In opam 2.0 (not released yet) the download tool verifies the TLS certificates~\cite{opampr} using the system-wide trustchain.
Nevertheless, a breach of the repository server enables an attacker to modify updates and tarballs arbitrarily.
The (known weak~\cite{rfc6151}) MD5 digest of source tarballs is stored in the repository as metadata.
Neither the build recipe nor additional files are checksummed.

In this paper we present \emph{conex}, a system which uses end-to-end digital signatures (the original author signs their package and build instructions, the client verifies this signature) to re-establish trust into opam.
Using end-to-end signature, the central server hosting the repository does not need to be trusted, neither does the server where the tarballs are downloaded from need to be trusted!
Conex and opam integrate well, and conex has been designed with opam in mind, but it can be used without opam to publish and update data in a way that it can be authenticated.
The main opam repository is a git repository hosted on GitHub.
While conex makes use of GitHub identifiers, it does not depend at all on GitHub, and not even on git.
Conex can also be used with other public or private opam repositories, only an initial set of trust anchors has to be distributed in a safe manner (e.g. an opam package in the main opam repository).
The main goal of conex is to provide a secure way to distribute packages.
Another design goal is to retain the workflow of package authors and opam users.
Janitors (repository maintainers) need to adapt to a slightly more complex workflow.

Conex is implemented in OCaml using only few dependencies (it should be installable without opam), and licensed under the 2-clause BSD license (source~\cite{conex}).

Earlier work on securing software update systems, such as the update framework~\cite{tuf} and diplomat~\cite{diplomat} focus on distributing directories securely, but make simultaneous updates of several packages by different authors hard, due to centralised files with authorisation and key information).
Our earlier proposal for opam signing~\cite{opamsigning} adjusted ideas from TUF to the current opam workflow, but introduced superfluous structure.
Conex still builds on top of tuf and diplomat, but simplifies our original proposal by introducing less structure and being less complex.

In section~\ref{sec:design} we will discuss the design of conex, followed by reasonable shortcuts in section~\ref{sec:shortcuts}.
Afterwards in section~\ref{sec:implementation} we present the implementation and opam integration.
We will discuss related in section~\ref{sec:related} and conclude with future work in section~\ref{sec:conclusion}.

\section{Design} \label{sec:design}
We first take a step back and do not consider all the details of opam for now, but focus on how to distribute a repository in a secure way.

Conex does not claim to protect against undiscovered vulnerabilities in the packages, dependent tools, or conex itself.
It is also not a scanner for installed vulnerable packages.

\subsection{Threat model}
There are many risks that users of software update systems face, ranging from interjecting traffic over weaknesses in TLS, weaknesses in the network infrastructure, compromising signing keys, stealing keys, to compromising the repository or signing infrastructure.

This leads us to consider a threat model where parts of the involved systems are compromised.  We assume that an attacker can:
\begin{itemize}
  \item Compromise the central server.
  \item Respond to user requests (acting as man-in-the-middle).
\end{itemize}

An attack is successful if the attacker can change the contents of a package that a user installs, or preventing a user from updating to the most recent version of a package.

We include signatures and keys in the same repository as the package metadata, this prevents attackers to hinder a user to receive key updates while updating metadata (possibly signed with a just compromised and revoked key).
We embed a monotonic increasing counter in each piece of data to prevent rollback attacks.

Compromise of (offline!) keys cannot be detected automatically, but requires janitors, authors, and other people to carefully read updates.
We assume that authors and janitors protect their keys reasonably.
If a key is compromised, we consider conex security to be effective if the impact is limited to the authorised packages of the compromised key.

Both a snapshot of the repository can be verified using a set of initial trust anchors (public keys of janitors), and an update (patch) to an existing verified repository which leads to a verified repository with the patch applied.

\subsection{Roles}
There are two roles in conex:
\begin{itemize}
  \item An \emph{author} who develops software and releases it into the repository.
  \item A \emph{janitor} who keeps the repository tidy and in a working state (fixing up reverse dependencies etc.).  Janitors also maintain the package name and author name resources.
\end{itemize}

An author signs their own packages, and janitors have to introduce authors and authorise authors for their packages.
A janitor has more privileges, and is also likely be an author of their packages.
In section~\ref{sec:shortcuts} we discuss some ways to make conex easier to use, leveraging the security level.

If a single janitor would be able to adjust packages which are missing version constraints (due to new releases of dependencies), revoke or update an author's key, and/or modify the set of janitors, then compromising a single janitor key would be sufficient to maliciously update the repository.
We avoid this by requiring a quorum of janitors to sign updates.

\subsection{Resources}
The repository contains several resources: package metadata (checksums, build recipes, dependencies), author (public key, id, role) and authorisation (relation between package name and author).
The role field distinguishes janitors and authors.

Each resource is only valid if it is signed by either one of the authorised key identifiers, or a quorum of janitors.
Each resource contains a monotonic counter to prevent rollback attacks.

Packages, identified by their name, and key identifiers, are the names we need to protect against starvation attacks.
Janitors need to faithfully verify new package and public key submissions.

Resource updates are distributed (each author updates their own packages), and should be free of conflicts (conflicts are detected during merge, it is up to the janitors to decide which update should go through, and which should be resubmitted on top of the other).

\subsection{Repository layout}
While conex is not bound to a file system, we'll reuse their terminology.
The root directory has two subdirectories, \texttt{keys} and \texttt{data}.

The \texttt{keys} directory contains one file, named after the key identifier.
The \texttt{data} directory consists of subdirectories, which each must contain a file \texttt{authorisation} and a file \texttt{index}, plus possibly any subdirectories.
Each subdirectory must be listed in the \texttt{index} file, and contains a \texttt{checksum} file besides any other data files and directories (all listed in the checksum file).

Each file contains a set of key-value pairs.
Any key-value representation which can be non-ambiguously normalised into a string, can be used (such as the opam format, json, s-expressions).
Each file contains a \texttt{signed} structure: \texttt{(signed: }\textbf{resource}\texttt{, sigs: }\textbf{(id, signature) list}\texttt{)}.
Resources, either a public key, an authorisation, or a checksum, have been mentioned before, and are explained in more detail below.
The signature is computed over the normalised string representation of the resource concatenated with the public key id (separated by a space character).

\begin{figure}
\textbf{id}: 7bit ASCII case sensitive string\\
\textbf{name}: 7bit ASCII case sensitive string\\
\textbf{signature}: Base64 encoded RSASSA-PSS with SHA256 (PKCS1)\\
\textbf{public key}: RSA public key ($\geq$ 2048 bit, PKCS1, PEM encoded)\\
\textbf{role}: \texttt{Janitor} $|$ \texttt{Author} $|$ \texttt{Other} of \textbf{id}\\
\textbf{service}: \texttt{EMail} of \textbf{name} $|$ \texttt{GitHub} of \textbf{name} $|$ \texttt{Other} of \textbf{id} * \textbf{name}\\
\textbf{digest}: SHA256 digest of file\\
\textbf{byte size}: size as \textbf{Int64} of file
\caption{\label{fig:datatypes0}
  Data types for conex version 0}
\end{figure}

The data types for version 0 of conex are presented in Figure~\ref{fig:datatypes0}.
Distinction between id and name is only for clarity (ids are the public key ids, where names are the package names), they form two disjoint sets of identifiers.
Upon insertion of new identifiers, checks for non-collision and validity need to be done.
The file formats of keys (Figure~\ref{fig:keyformat}), authorisations (Figure~\ref{fig:authformat}), index (Figure~\ref{fig:indexformat}), and checksums (Figure~\ref{fig:checksumformat}) include a version number (0 for this proposal) and a monotonic counter (starting at 0) each.
The index is signed by an authorised author, and contains a list of releases of the given package name.
The list of files in the checksum contains a list of all (recursively) present files in the current directory.
Files not mentioned in the checksum are not exposed from conex to opam.
Instead of removing author ids and packages, we put empty placeholders signed by janitors into the repository.
Otherwise an attacker can remove all the author ids and packages, since there's no signed trace of this change.

\begin{figure}
  \centering
  \begin{minipage}{\hsize}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
key: \textbf{public key},\\
identifier: \textbf{id},\\
accounts: \textbf{service list},\\
role: \textbf{role}
 \caption{\label{fig:keyformat} Public key}
    \end{minipage}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
name: \textbf{name},\\
authorised: \textbf{id list}
 \caption{\label{fig:authformat} Authorisation}
    \end{minipage}
  \end{minipage}
\end{figure}

\begin{figure}
  \centering
  \begin{minipage}{\hsize}
    \begin{minipage}{.3 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
name: \textbf{name},\\
releases: \textbf{name list}
 \caption{\label{fig:indexformat} Index}
    \end{minipage}
    \begin{minipage}{.55 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
name: \textbf{name},\\
files: \textbf{(name, byte size, digest) list}
 \caption{\label{fig:checksumformat} Checksum}
    \end{minipage}
  \end{minipage}
\end{figure}

\subsection{Chain of Trust}
A package is valid if and only if the chain of trust can be verified.
All files in the subdirectory are registered in the checksum file, which is signed by either an author (listed in the authorisation file in the parent directory) or a quorum of janitors.
The subdirectory name must be present in the valid index file of the parent directory.
The authorisation file itself has to be signed by a quorum of janitors (whose keys need to be in the repository and appropriately signed).
Public keys have to be signed by a quorum of janitors as well.

To bootstrap the chain, we assume an existing set of janitors (which cardinality is above the quorum).
The set for the main opam repository is distributed with opam, thus a compromise of opam is fatal (out of scope for conex, need to be authenticated via other means).

\subsection{Snapshot verification}
Verification of a snapshot is straightforward: take the set of initial janitors as trusted, then iterate through all packages, verify the authorisation file (properly signed by a quorum of janitors), and verify all checksum files.
The public keys are verified upon demand (whenever a signature is encountered, the key of the signing id is verified).

\subsection{Update verification}
Consider that the repository in the old state is verified, use it to verify the update (in form of a patch) in the following way:
\begin{itemize}
  \item key modification -- key is sufficiently big, unique id matching filename, counter is increased, signed by quorum, no signature with old key
  \item key insertion -- key is sufficiently big, unique id matching filename, counter is 0, signed by quorum
  \item key deletion -- counter increasing, no signatures with this key, empty string is used to denote no key, signed by quorum
  \item role modification -- counter increasing, no signature with old role, signed by quorum
  \item authorisation insertion -- counter is 0, id matches directory, signed by quorum, authorised key ids are present
  \item authorisation modification -- counter increasing, all checksum are signed with an authorised key, signed by quorum
  \item complete package removal -- empty authorised, empty releases, signed by quorum
  \item package insertion -- valid authorisation, index, and checksum, both counter 0, signed by author (authorisation by quorum)
  \item package update -- valid checksum, counter increasing, signed by author
  \item new package version release -- index counter increasing and new version added to list, valid checksum, signed by author
  \item package version deletion -- counter increasing, removed from releases, signed by author
\end{itemize}

\subsection{Security Analysis}
Assuming a man-in-the-middle-attacker between the repository and a user wants to deliver a backdoored package.
The first step is to insert a backdoor, which can either be done by modifying the source code of a package (either updating the source tarball, preparing a new release, or introducing a separate patch file) in the repository, or introducing a new package and a dependency.
All these mechanism require either the original author key (or a quorum of janitors).
Any suspicious behaviour in widely used packages will likely be spotted by the community which reads through the commit logs.

An attacker can freeze a client from updating (by preventing the client to communicate with the repository server), but downgrades of packages are not possible (an update cannot remove packages without appropriate signatures).

Mix and match (providing outdated versions of package A, newer of package B, up-to-date package C) are not mitigated in conex since there is no central list of releases.
In the next section, we will discuss a timestamping service which mitigates mix and match attacks (and let a user detect freeze attacks).

Conex does not depend on unreliable system time and comparison of timestamps, instead it uses monotonic counters.
It handles revocation explicitly by retaining the resource in the repository, but leaving the data element (e.g. public key) empty.
Other systems rely on git signing, which is out-of-tree data and thus an attacker can prevent a user from receiving this data (while still receiving (potentially malicious) updates).

In the case conex cannot verify package metadata, it will not pass this unverified data to opam.
In case conex cannot verify a downloaded tarball, this will be reported.

The weakest link of security are humans and their computer systems - janitors and authors will have their keys compromised after serving some website or inserting a USB stick.

\section{Shortcuts} \label{sec:shortcuts}
In the presented setup, conex requires verification of at least 8 signatures (3 for the public key, 3 for the authorisation, then index and checksum) per package.
Valid public key can be cached during processing, leaving the number at 5, which is still too high, especially for embedded systems.

In addition, the previously mentioned mix and match and freeze attack are possible.
To mitigate this, a \emph{timestamp service}, which updates the repository regularly, verifies all patches, and attaches its digital signature to the latest patch (via an git annotated tag over the git commit).
The public key of the timestamp service is inside the repository and signed by multiple janitors.
Clients can, instead of verifying packages and updates independently, trust the timestamp service (or multiple timestamp services).
This will reduce computational overhead (signature verification are expensive operations on big numbers), and might be worth for embedded devices.
If there are multiple timestamp services run by different organisations, and its code is well audited, it might be reasonable to use them as default.
While there is not yet an implementation of a timestamp service, it will likely be implemented as a MirageOS~\cite{mirage} application, thus the trusted code base is small.

The main opam repository served as an example over the last section, but conex can easily be applied to any other opam repository.
Initial set of janitors (can be a single element), the quorum (can be one) must be distributed for each other repository.
Of course, an automated timestamp service can be used for other repositories as well.
Conex can also securely distribute binary repositories (since OCaml 4.03 it emits reproducible binaries).

If the work overhead for janitors turns out too high, we can leverage the system in a way that \emph{neither authorisation nor public keys require a quorum}.
This will open the system to name squatting attacks, and in a snapshot authorisations cannot be verified that they contain the same information as the original ones, but updates can be verified (and again, we can outsource trust into the timestamping service).

Another useful automated system will be a \emph{PR bot}, similar to Camelus, which verifies each PR on the main opam repository and reports the results of the update (whether it is valid, how many janitors need to sign, etc.).

\section{Conex implementation} \label{sec:implementation}
An early version of conex~\cite{conex} is implemented in OCaml, but not yet released at the time of writing.
It does not depend on opam-lib, but instead will be functorised over the data layout (json, opam, s-expression).
To further minimise the dependencies (installed via opam), the goal is to have a cryptographic module which uses the widely deployed openssl utility for bootstrapping.

The timestamp service and other automated tools will use the conex library as a basis.

Conex is not yet integrated into opam at all (but some work on opam 2.0 will make integration straightforward).
The idea is to use conex as a separate application which proxies discover and download commands done by opam.

Janitors need more tooling support, especially to sign off quorums.
The idea is to have a git branch per update requiring a quorum, and the janitors adding their signatures to that branch as separate commits.
Once the quorum is reached, the branch can be squashed and merged into the repository.
A queue of updates needing attention (a quorum) will be provided by the tooling.

\section{Related work} \label{sec:related}
Conex is based on the update framework~\cite{tuf}, which centralises all packages into a single \texttt{target.txt} file, and has more structure in keys.
Instead of doing this, conex was designed with version control and distributed updates in mind.

A more recent contribution is Diplomat~\cite{diplomat}, but instead of automatically promoting unclaimed packages to claimed (which we can do using timestamp services), we believe in end-to-end signing of packages.
Our security model relies on offline keys which are protected by authors.

The Haskell signing proposal and implementation does not enforce authors to sign their packages, but instead the repository is signed.
Thus, the repository server (or key server) needs to be trusted ultimately.

\section{Conclusion} \label{sec:conclusion}
We presented conex: securing opam, a software update system, without modifying the current workflows too much.
The security will be based on offline keys and end-to-end signing of packages.
This removes any central server from the trust chain.
Attacks, observed in practise, and theoretic from related work, are mitigated by conex.

Future work includes finishing conex, writing more tests, integration into opam, and develop automated services.

\bibliographystyle{acm}
\bibliography{paper}

\end{document}

### 2016-06-08 paper.tex
.

@@
 
-\title{Conex - establishing trust into repositories}
+\title{Conex - establishing trust into data repositories}
 
@@
 \begin{abstract}
-  Software update systems, such as opam, should verify that packages and metadata thereof are authenticated.
-  Opam does not include any mechanism to verify authenticity at the moment.
-  We propose \emph{conex}, which uses end-to-end digital signatures from the package author to the user.
-  Conex is implemented in OCaml and can be integrated into opam to build a security story where even the central repository server does not need to be trusted.
-  Authors and users will not need to change their workflow, whereas janitors (repository maintainers) will have to adjust their workflow.
+  Opam is a software update systems, responsible for discovering, downloading, building, and installing packages.
+  The data and metadata at each step should be authenticated to originate from a (transitively) trusted author.
+  Opam does not include any mechanism to authenticate any step at the moment.
+  We propose \emph{conex}, which establishes end-to-end digital signatures from the package author to the user.
+  Using conex, neither the distribution server nor the transport layer need to be trusted.
+  Authors and users need to adapt their workflow only slightly (or use opam-publish), whereas janitors (repository maintainers) need to use more sophisticated tools.
 \end{abstract}
@@
 \section{Introduction}
-Software update systems, such as opam, need to discover, download, and install packages.
-Crucial to all steps is that the received data is authenticated before being used to prevent attacks injecting backdoors.
-A software update system relies on a central server which provides a repository (source code and metadata such as dependencies and build recipe) of all packages.
-When a new package is released, the repository is updated.
-The central server is an online services and a single point of failures, and recently these servers been attacked \cite{adobe,apache2,apache1,opera,kernel,cabal,kernelorg,debian,freebsd,php3,savannah,maven,savannah2,npm,php,phpmyadmin,php2,github,ocamlforge,redhatssh,gentoo,rubygems,fedora,wordpress} to distribute malware.
+Opam is a software update systems, and needs to discover, download, build, and install packages.
+To prevent injection of malware, all steps need to authenticate the received data before processing it.
+A software update system relies on a central server which distributes a repository (source code and metadata such as dependencies and build recipe) of all packages.
+When a new package is released, it is inserted into the repository and distributed to all clients by the central server.
+The central server is an online service and a single point of failure, and recently these central servers have been attacked \cite{adobe,apache2,apache1,opera,kernel,cabal,kernelorg,debian,freebsd,php3,savannah,maven,savannah2,npm,php,phpmyadmin,php2,github,ocamlforge,redhatssh,gentoo,rubygems,fedora,wordpress} to distribute packages with injected malware.
 
-While so far there is no known breach of opam installations (but other package managers have been compromised, see \cite{cabal,maven,npm,rubygems}), there is not much security in place:
-In opam 1.2.2 none of the steps (discover, download, install) includes verification.
-The discover and download steps use https, but pass \texttt{--no-check-certificates} to the download utility \texttt{wget}, which disables certificate verification.
-Any man-in-the-middle between client and repository can arbitrarily modify updates or tarballs.
-The build and install steps may do any modifications to the file system, not limited to the build and target directory of the package.
-In opam 2.0 (not released yet) the download tool verifies the TLS certificates~\cite{opampr} using the system-wide trustchain.
-Nevertheless, a breach of the repository server enables an attacker to modify updates and tarballs arbitrarily.
-The (known weak~\cite{rfc6151}) MD5 digest of source tarballs is stored in the repository as metadata.
-Neither the build recipe nor additional files are checksummed.
+So far there has been no breach of opam detected, but other package managers have been compromised, see \cite{cabal,maven,npm,rubygems}.
+Opam does not provide any security mechanism at any step:
+Discovery in current opam (1.2.2) is done via a https download (for the main opam repository), but the download tool (\texttt{wget}) is used with the \texttt{--no-check-certificates} command line option, disabling any certificate validation.
+Thus, a person-in-the-middle can inject arbitrary metadata (additional patches, modified packages, modified checksums, ...) during the discovery step.
+A package is downloaded using the url specified in the metadata of the package (usually https, again without certificate validation), and checks the recorded MD5 digest (which is known to be weak~\cite{rfc6151}) against the downloaded archive.
+The build step is not contained into a temporary chroot environment, but may modify arbitrary files on disk.
+The recipe for building is part of the metadata, and may also include invoking any shell commands (such as downloading more things).
+The install step may also modify arbitrary files.
 
-In this paper we present \emph{conex}, a system which uses end-to-end digital signatures (the original author signs their package and build instructions, the client verifies this signature) to re-establish trust into opam.
-Using end-to-end signature, the central server hosting the repository does not need to be trusted, neither does the server where the tarballs are downloaded from need to be trusted!
-Conex and opam integrate well, and conex has been designed with opam in mind, but it can be used without opam to publish and update data in a way that it can be authenticated.
-The main opam repository is a git repository hosted on GitHub.
-While conex makes use of GitHub identifiers, it does not depend at all on GitHub, and not even on git.
-Conex can also be used with other public or private opam repositories, only an initial set of trust anchors has to be distributed in a safe manner (e.g. an opam package in the main opam repository).
-The main goal of conex is to provide a secure way to distribute packages.
-Another design goal is to retain the workflow of package authors and opam users.
-Janitors (repository maintainers) need to adapt to a slightly more complex workflow.
+Opam 2.0 is not released yet, but fixes several major concerns:
+downloading now validates the TLS certificates~\cite{opampr} using the system-wide trustchain,
+a wrap command is available for building and installing packages~\cite{wrappr}.
+Nevertheless, a breach of the central server enables an attacker to modify packages arbitrarily.
+Relying only on transport layer security for packages, where many implementations have a long history of security vulnerabilities (e.g. Apple's goto fail), is not optimal.
 
-Conex is implemented in OCaml using only few dependencies (it should be installable without opam), and licensed under the 2-clause BSD license (source~\cite{conex}).
+In this paper we present \emph{conex}, a system which uses end-to-end digital signatures - the original author signs their package and build instructions, the client verifies this signature - to re-establish trust into packages installed by opam.
+Using end-to-end signatures, neither the the server hosting the repository nor the server where the tarballs are downloaded from need to be trusted!
+Conex has been designed with opam in mind, but it can be used without opam to publish and update data in an authenticated way.
+At the moment, the main opam repository is a git repository hosted on GitHub.
+While conex uses GitHub identifiers, it does not depend at all on GitHub (could use EMail addresses instead), and not even on git.
+Conex is not limited to the main opam repository, but can be used with any public or private opam repository, or any other data repository.
+The only requirement is a second channel to distribute an initial set of trust anchors (e.g. an opam package in the main opam repository).
+The main goal of conex is to \emph{establish an authenticated way to distribute packages}.
+Conex is constrained by the current use of the opam repository:  most package authors upload their releases as a pull request, janitors (repository maintainers) merge these after manual and automated (Camelius) feedback (on package name, missing dependencies, build-only dependencies, ...).
+Using conex, most of the burden has to be carried by the janitors.
 
-Earlier work on securing software update systems, such as the update framework~\cite{tuf} and diplomat~\cite{diplomat} focus on distributing directories securely, but make simultaneous updates of several packages by different authors hard, due to centralised files with authorisation and key information).
+Conex is implemented in OCaml using only dependencies of opam (such as cmdliner), and licensed under the 2-clause BSD license (source~\cite{conex}).
+It should be installable at the same time as opam itself to provide authentication from the start.
+
+Earlier work on securing software update systems, such as the update framework (TUF)~\cite{tuf} and Diplomat~\cite{diplomat} focus on distributing directories securely, but make simultaneous updates of several packages by different authors hard, due to centralised files with authorisation and key information.
 Our earlier proposal for opam signing~\cite{opamsigning} adjusted ideas from TUF to the current opam workflow, but introduced superfluous structure.
-Conex still builds on top of tuf and diplomat, but simplifies our original proposal by introducing less structure and being less complex.
 
@@
 \section{Design} \label{sec:design}
-We first take a step back and do not consider all the details of opam for now, but focus on how to distribute a repository in a secure way.
+We first take a step back and do not consider all the details of opam for now.
 
@@
 \begin{itemize}
-  \item Compromise the central server.
-  \item Respond to user requests (acting as man-in-the-middle).
+  \item Compromise the central server distributing the repository.
+  \item Respond to user requests (acting as person-in-the-middle).
 \end{itemize}
@@
 
-We include signatures and keys in the same repository as the package metadata, this prevents attackers to hinder a user to receive key updates while updating metadata (possibly signed with a just compromised and revoked key).
-We embed a monotonic increasing counter in each piece of data to prevent rollback attacks.
-
-Compromise of (offline!) keys cannot be detected automatically, but requires janitors, authors, and other people to carefully read updates.
-We assume that authors and janitors protect their keys reasonably.
-If a key is compromised, we consider conex security to be effective if the impact is limited to the authorised packages of the compromised key.
-
-Both a snapshot of the repository can be verified using a set of initial trust anchors (public keys of janitors), and an update (patch) to an existing verified repository which leads to a verified repository with the patch applied.
+We cannot detect compromises of (offline!) keys, but this requires janitors, authors, and other people to carefully read through updates.
+We assume that authors and janitors protect their private keys on their computers in a reasonable way, but we also have mechanisms in place for key revocation.
+If a private key is compromised, we consider conex security to be effective if the impact is limited to the authorised packages of the compromised key.
 
@@
 \begin{itemize}
-  \item An \emph{author} who develops software and releases it into the repository.
+  \item An \emph{author} who develops packages and releases them to the repository.
   \item A \emph{janitor} who keeps the repository tidy and in a working state (fixing up reverse dependencies etc.).  Janitors also maintain the package name and author name resources.
@@
 
-An author signs their own packages, and janitors have to introduce authors and authorise authors for their packages.
-A janitor has more privileges, and is also likely be an author of their packages.
-In section~\ref{sec:shortcuts} we discuss some ways to make conex easier to use, leveraging the security level.
+An author signs their own packages.
+A janitor has to introduce authors, authorise authors for their packages, and also adjust packages due to updated dependencies (such as the version constraints, or minor patches to build on another platform).
+It is likely that a janitor also authors some packages, there is no need for multiple keys in this case.
+In section~\ref{sec:shortcuts} we discuss ways to make conex easier to use, leveraging the security level.
 
-If a single janitor would be able to adjust packages which are missing version constraints (due to new releases of dependencies), revoke or update an author's key, and/or modify the set of janitors, then compromising a single janitor key would be sufficient to maliciously update the repository.
-We avoid this by requiring a quorum of janitors to sign updates.
+If a single janitor would be trusted to tidy up packages, compromising only one janitor key would be sufficient to undermine the security of conex.
+We avoid this by requiring a quorum of janitors to sign hotfixes.
 
 \subsection{Resources}
-The repository contains several resources: package metadata (checksums, build recipes, dependencies), author (public key, id, role) and authorisation (relation between package name and author).
+We distinguish various resources which are all hosted in the same repository:
+package metadata (checksums, build recipe, dependencies), author information (public key, id, role), list of releases (of a single package), and authorisation (relation between package name and author).
 The role field distinguishes janitors and authors.
+For each janitor the repository contains a resource with all resources and checksums, signed by the janitor (to avoid way too many public key operations during sign and verify).
 
@@
 
-Packages, identified by their name, and key identifiers, are the names we need to protect against starvation attacks.
+There are two universes of names which we need to protect against squatting attacks (including typographics ones~\cite{typo}): package names and key identifiers.
 Janitors need to faithfully verify new package and public key submissions.
 
-Resource updates are distributed (each author updates their own packages), and should be free of conflicts (conflicts are detected during merge, it is up to the janitors to decide which update should go through, and which should be resubmitted on top of the other).
+Each author updates their own packages, which should be free of conflicts due to separate resources for each package.
+If a conflict is detected, it is up to the janitors to decide which update should go through, and which should be resubmitted on top of the other.
 
@@
 While conex is not bound to a file system, we'll reuse their terminology.
-The root directory has two subdirectories, \texttt{keys} and \texttt{data}.
+The root directory has three subdirectories, \texttt{keys}, \texttt{sigs} and \texttt{data}.
 
-The \texttt{keys} directory contains one file, named after the key identifier.
-The \texttt{data} directory consists of subdirectories, which each must contain a file \texttt{authorisation} and a file \texttt{index}, plus possibly any subdirectories.
-Each subdirectory must be listed in the \texttt{index} file, and contains a \texttt{checksum} file besides any other data files and directories (all listed in the checksum file).
+The \texttt{keys} directory contains one file for each author, named after the key identifier.
+The \texttt{sigs} directory contains one file for each janitor, named after the janitor key identifier.
+The \texttt{data} directory consists of subdirectories, which each must contain a file \texttt{authorisation} and a file \texttt{releases}, plus possibly any subdirectories.
+Each subdirectory must be listed in the \texttt{releases} file, and must contain a \texttt{checksum} file besides any other data files and directories (all listed in the checksum file).
 
@@
 Each file contains a \texttt{signed} structure: \texttt{(signed: }\textbf{resource}\texttt{, sigs: }\textbf{(id, signature) list}\texttt{)}.
-Resources, either a public key, an authorisation, or a checksum, have been mentioned before, and are explained in more detail below.
-The signature is computed over the normalised string representation of the resource concatenated with the public key id (separated by a space character).
+The different resources are explained in more detail below.
+The signature is computed over the normalised string representation of the resource concatenated with the public key id and the resource kind, separated by a space character.
 
 \begin{figure}
-\textbf{id}: 7bit ASCII case sensitive string\\
-\textbf{name}: 7bit ASCII case sensitive string\\
+\textbf{id}: restricted string\\
+\textbf{name}: restricted string\\
 \textbf{signature}: Base64 encoded RSASSA-PSS with SHA256 (PKCS1)\\
@@
 The data types for version 0 of conex are presented in Figure~\ref{fig:datatypes0}.
-Distinction between id and name is only for clarity (ids are the public key ids, where names are the package names), they form two disjoint sets of identifiers.
+A \texttt{restricted string} is a case insensitive 7 bit ASCII string without control characters or special characters.
+Depending on usage, more characters are forbidden (e.g. opam package names may not contain a '.').
+Distinction of id and name is only for clarity (ids are the public key ids, where names are the package names), they form two disjoint sets of identifiers.
 Upon insertion of new identifiers, checks for non-collision and validity need to be done.
-The file formats of keys (Figure~\ref{fig:keyformat}), authorisations (Figure~\ref{fig:authformat}), index (Figure~\ref{fig:indexformat}), and checksums (Figure~\ref{fig:checksumformat}) include a version number (0 for this proposal) and a monotonic counter (starting at 0) each.
-The index is signed by an authorised author, and contains a list of releases of the given package name.
-The list of files in the checksum contains a list of all (recursively) present files in the current directory.
-Files not mentioned in the checksum are not exposed from conex to opam.
-Instead of removing author ids and packages, we put empty placeholders signed by janitors into the repository.
-Otherwise an attacker can remove all the author ids and packages, since there's no signed trace of this change.
+
+Each file includes a data version (0 in this proposal) and a monotonic 64 bit counter (starting at 0).
 
@@
 
+The file format for a public key, presented in Figure~\ref{fig:keyformat}, contains the PEM encoded public key, the identifier (used for signatures and authorisations in opam), accounts in other systems, and its role.
+
+The authorisations file format, shown in Figure~\ref{fig:authformat}, contains the package name and a list of authorised key identifiers for this package.
+
 \begin{figure}
@@
 releases: \textbf{name list}
- \caption{\label{fig:indexformat} Index}
+ \caption{\label{fig:releasesformat} Releases}
     \end{minipage}
@@
 
+The releases file format, shown in Figure~\ref{fig:releasesformat}, contains the package name and the list of all releases of this package.
+It is signed by an authorised author (or a quorum of janitors).
+This is crucial to avoid rollback attacks, where an attacker hinders a client from receiving package updates.
+
+The checksums file format, shown in Figure~\ref{fig:checksumformat}, contains the release name, and a list of file names which are part of the release (opam, url, descr, optionally patches, archive) and their sizes and digests.
+This list must contain all files recursively present in the current directory (excluding the checksum file itself).
+
+\begin{figure}
+  \centering
+counter: \textbf{Int64},\\
+version: \textbf{Int64},\\
+identifier: \textbf{id},\\
+resources: \textbf{(name, kind, digest) list}
+ \caption{\label{fig:janindexformat} Janitor index}
+\end{figure}
+
+The janitor index file format, shown in Figure~\ref{fig:janindexformat}, contains the janitor identifier, and all resources the janitor vouches for.
+
+Instead of removing author ids and packages, we put empty placeholders signed by janitors into the repository.
+
 \subsection{Chain of Trust}
@@
 All files in the subdirectory are registered in the checksum file, which is signed by either an author (listed in the authorisation file in the parent directory) or a quorum of janitors.
-The subdirectory name must be present in the valid index file of the parent directory.
-The authorisation file itself has to be signed by a quorum of janitors (whose keys need to be in the repository and appropriately signed).
+The subdirectory name must be present in the valid releases file of the parent directory.
+The authorisation file itself has to be signed by a quorum of janitors.
 Public keys have to be signed by a quorum of janitors as well.
@@
 To bootstrap the chain, we assume an existing set of janitors (which cardinality is above the quorum).
-The set for the main opam repository is distributed with opam, thus a compromise of opam is fatal (out of scope for conex, need to be authenticated via other means).
+The set for the main opam repository is distributed with opam.
+The initial authenticated opam installation, including the initial set of janitors, is out of scope for conex (requires another package manager, OpenPGP signatures, ...).
 
-\subsection{Snapshot verification}
-Verification of a snapshot is straightforward: take the set of initial janitors as trusted, then iterate through all packages, verify the authorisation file (properly signed by a quorum of janitors), and verify all checksum files.
-The public keys are verified upon demand (whenever a signature is encountered, the key of the signing id is verified).
+\subsection{Verification}
+Both a snapshot of the repository can be verified using a set of initial trust anchors (public keys of janitors), and an update (patch) to an existing verified repository which leads to a verified repository with the patch applied.
 
-\subsection{Update verification}
-Consider that the repository in the old state is verified, use it to verify the update (in form of a patch) in the following way:
+\paragraph{Snapshot verification}  Verification of a snapshot is straightforward: take the set of initial janitors as trusted, then iterate through all packages, verify the authorisation file (properly signed by a quorum of janitors), all releases files, and verify all checksum files.
+The public keys are verified upon demand: whenever a signature is encountered, the key of the signing id is verified.
+
+\paragraph{Update verification}  Consider that the repository in the old state is verified, use it to verify the update (in form of a patch) in the following way:
 \begin{itemize}
@@
   \item complete package removal -- empty authorised, empty releases, signed by quorum
-  \item package insertion -- valid authorisation, index, and checksum, both counter 0, signed by author (authorisation by quorum)
+  \item package insertion -- valid authorisation, releases, and checksum, both counter 0, signed by author (authorisation by quorum)
   \item package update -- valid checksum, counter increasing, signed by author
@@
 \subsection{Security Analysis}
-Assuming a man-in-the-middle-attacker between the repository and a user wants to deliver a backdoored package.
-The first step is to insert a backdoor, which can either be done by modifying the source code of a package (either updating the source tarball, preparing a new release, or introducing a separate patch file) in the repository, or introducing a new package and a dependency.
-All these mechanism require either the original author key (or a quorum of janitors).
+%\TODO{from opam signing: When a developer key is stolen, the snapshot key is also needed to target a single user for a malicious patch.}
+%\TODO{still sort of a laundry list}
+
+Package metadata, signatures, and public keys are distributed in the same repository to prevent attackers from hindering one communication link, e.g. the public key one.
+This would otherwise allow a detected attacker who compromised a key to deliver metadata signed with this compromised key.
+
+Each resource includes a monotonically increasing counter to prevent rollback attacks.
+Otherwise, an attacker could replay e.g. an insecure package build recipe.
+Also, each resource includes the name, otherwise the signature of package A would be valid for package B.
+Each signature contains the resource kind to avoid carrying over a signature from one resource to another.
+
+Instead of removing resources to revoke them, the already claimed names stay there.
+The public key, package, or janitor index resource is emptied and must be signed by a quorum of janitors.
+
+A person-in-the-middle attacker could implement a backdoor for a package, but would need to either modify the checksum, prepare a new release, embed the patch as metadata in the repository, or introduce a new package (which they could) and add a dependency.
+All of these actions require either a signature from the original author or from a quorum of janitors.
 Any suspicious behaviour in widely used packages will likely be spotted by the community which reads through the commit logs.
@@
 
-Mix and match (providing outdated versions of package A, newer of package B, up-to-date package C) are not mitigated in conex since there is no central list of releases.
-In the next section, we will discuss a timestamping service which mitigates mix and match attacks (and let a user detect freeze attacks).
-
 Conex does not depend on unreliable system time and comparison of timestamps, instead it uses monotonic counters.
-It handles revocation explicitly by retaining the resource in the repository, but leaving the data element (e.g. public key) empty.
 Other systems rely on git signing, which is out-of-tree data and thus an attacker can prevent a user from receiving this data (while still receiving (potentially malicious) updates).
@@
 
-The weakest link of security are humans and their computer systems - janitors and authors will have their keys compromised after serving some website or inserting a USB stick.
+Mix-and-match attacks provide e.g. outdated releases of package A and B, up-to-date releases of package C, are not mitigated in conex since there is no central list of releases.
+They can be mitigated by relying on some append-only log, blockchain, or other monotonic updates, such as git without forced pushes.
+We will discuss a timestamp notary in the next section, mitigating both mix-and-match and freeze attacks.
 
 \section{Shortcuts} \label{sec:shortcuts}
-In the presented setup, conex requires verification of at least 8 signatures (3 for the public key, 3 for the authorisation, then index and checksum) per package.
-Valid public key can be cached during processing, leaving the number at 5, which is still too high, especially for embedded systems.
+%\TODO{see also http://bentnib.org/posts/2016-04-12-authenticated-data-structures-as-a-library.html}
 
-In addition, the previously mentioned mix and match and freeze attack are possible.
-To mitigate this, a \emph{timestamp service}, which updates the repository regularly, verifies all patches, and attaches its digital signature to the latest patch (via an git annotated tag over the git commit).
+%In the presented setup, conex requires verification of at least 8 signatures (3 for the public key, 3 for the authorisation, then index and checksum) per package.
+%Valid public key can be cached during processing, leaving the number at 5, which is still too high, especially for embedded systems.
+
+To mitigate mix-and-match attacks, a \emph{timestamp notary}, which updates the repository regularly, verifies all updates, and attaches its digital signature to the latest patch (via an git annotated tag over the git commit).
 The public key of the timestamp service is inside the repository and signed by multiple janitors.
@@
 This will reduce computational overhead (signature verification are expensive operations on big numbers), and might be worth for embedded devices.
-If there are multiple timestamp services run by different organisations, and its code is well audited, it might be reasonable to use them as default.
-While there is not yet an implementation of a timestamp service, it will likely be implemented as a MirageOS~\cite{mirage} application, thus the trusted code base is small.
+If there are multiple timestamp notaries run by different organisations, and its code is well audited, it might be reasonable to use them as default.
+While there is not yet an implementation of a timestamp notary, it will likely be implemented as a MirageOS~\cite{mirage,nqsb-tls} application, thus the trusted code base is small.
 
@@
 Initial set of janitors (can be a single element), the quorum (can be one) must be distributed for each other repository.
-Of course, an automated timestamp service can be used for other repositories as well.
-Conex can also securely distribute binary repositories (since OCaml 4.03 it emits reproducible binaries).
+Of course, an automated timestamp notary can be used for other repositories as well.
 
+%\TODO{this is insecure, since an attacker can add themselves to the authorisation file and self-sign it!}
 If the work overhead for janitors turns out too high, we can leverage the system in a way that \emph{neither authorisation nor public keys require a quorum}.
-This will open the system to name squatting attacks, and in a snapshot authorisations cannot be verified that they contain the same information as the original ones, but updates can be verified (and again, we can outsource trust into the timestamping service).
+This will open the system to name squatting attacks for identifiers.
+In a snapshot authorisations cannot be verified anymore, because there is no history to the authorisation.
+Update verification still applies: if we trust the timestamp notaries and use git, janitors don't have to sign public keys and authorisation files.
+
+%that they contain the same information as the original ones, but updates can be verified (and again, we can outsource trust into the timestamping notary).
 
@@
 
+Automated systems which build a package and record the (platform-dependent) digests of the resulting binaries can be integrated to sign their results in the repository.
+Since OCaml 4.03 produces reproducible binaries, thus the locally produced binaries can be checked.
+Conex can also serve for binary distribution.
+
 \section{Conex implementation} \label{sec:implementation}
-An early version of conex~\cite{conex} is implemented in OCaml, but not yet released at the time of writing.
+%\TODO{UX: timestamp in signature!}
+
+Conex is work-in-progress~\cite{conex}.
 It does not depend on opam-lib, but instead will be functorised over the data layout (json, opam, s-expression).
-To further minimise the dependencies (installed via opam), the goal is to have a cryptographic module which uses the widely deployed openssl utility for bootstrapping.
+To further minimise the dependencies, the goal is to have a cryptographic module which uses the widely deployed \emph{openssl} utility for bootstrapping.
 
-The timestamp service and other automated tools will use the conex library as a basis.
+The timestamp notary and other automated tools will use the conex library as a basis.
 
-Conex is not yet integrated into opam at all (but some work on opam 2.0 will make integration straightforward).
+Conex is not yet integrated into opam, but some work on opam 2.0, such as compilers-as-packages, will make the integration smooth.
 The idea is to use conex as a separate application which proxies discover and download commands done by opam.
 
-Janitors need more tooling support, especially to sign off quorums.
-The idea is to have a git branch per update requiring a quorum, and the janitors adding their signatures to that branch as separate commits.
-Once the quorum is reached, the branch can be squashed and merged into the repository.
+There is still demand for more tooling support for janitors.
+In order to reach a quorum, individual janitors have to add the digest to their index file.
+Once the quorum is reached, the proposed PR can be merged.
 A queue of updates needing attention (a quorum) will be provided by the tooling.
@@
 
+Our original opam signing proposal did not include a list of releases per package, opening up to attacks where the user does not receive a complete list of releases.
+Additionally, the original proposal introduced an unneccessary hierarchy of keys, and complicated the setup.
+It relied on the timestamping notary or to use git as update protocol with appropriate protection of the transport layer.
+
 \section{Conclusion} \label{sec:conclusion}
 We presented conex: securing opam, a software update system, without modifying the current workflows too much.
-The security will be based on offline keys and end-to-end signing of packages.
+The security is based on offline keys and end-to-end signing of packages.
 This removes any central server from the trust chain.
-Attacks, observed in practise, and theoretic from related work, are mitigated by conex.
+Attacks, observed in practise, and theoretic ones in related work, are mitigated by conex.
 

### 2016-06-09 paper.tex
flyspell

@@
 Thus, a person-in-the-middle can inject arbitrary metadata (additional patches, modified packages, modified checksums, ...) during the discovery step.
-A package is downloaded using the url specified in the metadata of the package (usually https, again without certificate validation), and checks the recorded MD5 digest (which is known to be weak~\cite{rfc6151}) against the downloaded archive.
+A package is downloaded using the URL specified in the metadata of the package (usually https, again without certificate validation), and checks the recorded MD5 digest (which is known to be weak~\cite{rfc6151}) against the downloaded archive.
 The build step is not contained into a temporary chroot environment, but may modify arbitrary files on disk.
@@
 Opam 2.0 is not released yet, but fixes several major concerns:
-downloading now validates the TLS certificates~\cite{opampr} using the system-wide trustchain,
+downloading now validates the TLS certificates~\cite{opampr} using the system-wide trust anchors,
 a wrap command is available for building and installing packages~\cite{wrappr}.
 Nevertheless, a breach of the central server enables an attacker to modify packages arbitrarily.
-Relying only on transport layer security for packages, where many implementations have a long history of security vulnerabilities (e.g. Apple's goto fail), is not optimal.
+Relying only on transport layer security for packages, where many implementations have a long history of security vulnerabilities (e.g. goto fail), is not optimal.
 
@@
 The main goal of conex is to \emph{establish an authenticated way to distribute packages}.
-Conex is constrained by the current use of the opam repository:  most package authors upload their releases as a pull request, janitors (repository maintainers) merge these after manual and automated (Camelius) feedback (on package name, missing dependencies, build-only dependencies, ...).
+Conex is constrained by the current use of the opam repository:  most package authors upload their releases as a pull request, janitors (repository maintainers) merge these after manual and automated (Camelus) feedback (on package name, missing dependencies, build-only dependencies, ...).
 Using conex, most of the burden has to be carried by the janitors.
@@
 
-There are two universes of names which we need to protect against squatting attacks (including typographics ones~\cite{typo}): package names and key identifiers.
+There are two universes of names which we need to protect against squatting attacks (including typographic ones~\cite{typo}): package names and key identifiers.
 Janitors need to faithfully verify new package and public key submissions.
@@
 Our original opam signing proposal did not include a list of releases per package, opening up to attacks where the user does not receive a complete list of releases.
-Additionally, the original proposal introduced an unneccessary hierarchy of keys, and complicated the setup.
+Additionally, the original proposal introduced an unnecessary hierarchy of keys, and complicated the setup.
 It relied on the timestamping notary or to use git as update protocol with appropriate protection of the transport layer.
@@
 \section{Conclusion} \label{sec:conclusion}
-We presented conex: securing opam, a software update system, without modifying the current workflows too much.
-The security is based on offline keys and end-to-end signing of packages.
-This removes any central server from the trust chain.
-Attacks, observed in practise, and theoretic ones in related work, are mitigated by conex.
+We presented conex, which secures the software update system opam with minor modifications to the current workflow.
+Conex uses end-to-end signatures (from package authors to users) as trust model.
+It removes any trust into a central server.
+Known attacks, observed in practise or described in literaturre, are mitigated sufficiently by conex.
 
-Future work includes finishing conex, writing more tests, integration into opam, and develop automated services.
+Future work includes finishing the implementation of conex, integrating it into opam, and develop automated services (timestamp notary, build bot, PR bot).
 

### 2016-06-09 paper.tex
Raphaël

@@
 At the moment, the main opam repository is a git repository hosted on GitHub.
-While conex uses GitHub identifiers, it does not depend at all on GitHub (could use EMail addresses instead), and not even on git.
+Conex, as presented here, uses GitHub identifiers, but it can use email addresses instead.
+Conex does not depend on GitHub, neither on git.
 Conex is not limited to the main opam repository, but can be used with any public or private opam repository, or any other data repository.
@@
 The main goal of conex is to \emph{establish an authenticated way to distribute packages}.
-Conex is constrained by the current use of the opam repository:  most package authors upload their releases as a pull request, janitors (repository maintainers) merge these after manual and automated (Camelus) feedback (on package name, missing dependencies, build-only dependencies, ...).
+Conex is constrained by the current use of the opam repository:  most package authors release a package and submit a pull request (PR) to the main opam repository on GitHub.
+Afterwards, janitors (repository maintainers) check the PR (good package name, potentially missing dependencies or build-time only dependencies), automated builds are performed by TravisCI, and the automated bot Camelus~\cite{camelus} reports feedback and missing fields.
 Using conex, most of the burden has to be carried by the janitors.
 
-Conex is implemented in OCaml using only dependencies of opam (such as cmdliner), and licensed under the 2-clause BSD license (source~\cite{conex}).
-It should be installable at the same time as opam itself to provide authentication from the start.
+Conex is implemented in OCaml and licensed under the 2-clause BSD license (source~\cite{conex}).
+To provide authenticity from the start, conex should be installed at the same time as opam, thus it is constrained to use few dependencies, and will only depend on cmdliner and opam-lib.
 
@@
 Afterwards in section~\ref{sec:implementation} we present the implementation and opam integration.
-We will discuss related in section~\ref{sec:related} and conclude with future work in section~\ref{sec:conclusion}.
+We will discuss related work in section~\ref{sec:related} and conclude with future work in section~\ref{sec:conclusion}.
 
@@
   \item An \emph{author} who develops packages and releases them to the repository.
-  \item A \emph{janitor} who keeps the repository tidy and in a working state (fixing up reverse dependencies etc.).  Janitors also maintain the package name and author name resources.
+  \item A \emph{janitor} who keeps the repository tidy and in a working state (fix up reverse dependencies etc.).  Janitors also maintain the package name and author name resources.
 \end{itemize}
@@
 An author signs their own packages.
-A janitor has to introduce authors, authorise authors for their packages, and also adjust packages due to updated dependencies (such as the version constraints, or minor patches to build on another platform).
-It is likely that a janitor also authors some packages, there is no need for multiple keys in this case.
+A janitor introduces authors, authorise authors for their packages, and also adjust packages due to updated dependencies (such as the version constraints, or minor patches to build on another platform).
+It is likely that a janitor acts as an author for their packages, there is no need for a janitor to have multiple private keys.
 In section~\ref{sec:shortcuts} we discuss ways to make conex easier to use, leveraging the security level.
@@
 package metadata (checksums, build recipe, dependencies), author information (public key, id, role), list of releases (of a single package), and authorisation (relation between package name and author).
-The role field distinguishes janitors and authors.
+The distinction between an author and a janitor is done via the role of the author information.
 For each janitor the repository contains a resource with all resources and checksums, signed by the janitor (to avoid way too many public key operations during sign and verify).
@@
 
-Our original opam signing proposal did not include a list of releases per package, opening up to attacks where the user does not receive a complete list of releases.
+Our original opam signing proposal~\cite{opamsigning} did not include a list of releases per package, opening up to attacks where the user does not receive a complete list of releases.
 Additionally, the original proposal introduced an unnecessary hierarchy of keys, and complicated the setup.

### 2016-06-09 paper.tex
condense

@@
 \section{Introduction}
-Opam is a software update systems, and needs to discover, download, build, and install packages.
-To prevent injection of malware, all steps need to authenticate the received data before processing it.
-A software update system relies on a central server which distributes a repository (source code and metadata such as dependencies and build recipe) of all packages.
+Opam discovers, downloads, builds, and installs OCaml packages.
+To prevent injection of malware, all received data needs to be authenticated.
+A software update system relies on a central server which distributes a repository (package metadata: dependencies and build instructions) of packages.
 When a new package is released, it is inserted into the repository and distributed to all clients by the central server.
-The central server is an online service and a single point of failure, and recently these central servers have been attacked \cite{adobe,apache2,apache1,opera,kernel,cabal,kernelorg,debian,freebsd,php3,savannah,maven,savannah2,npm,php,phpmyadmin,php2,github,ocamlforge,redhatssh,gentoo,rubygems,fedora,wordpress} to distribute packages with injected malware.
+The central server is an online service and a single point of failure.
+Recently those servers have been under attack \cite{adobe,apache2,apache1,opera,kernel,cabal,kernelorg,debian,freebsd,php3,savannah,maven,savannah2,npm,php,phpmyadmin,php2,github,ocamlforge,redhatssh,gentoo,rubygems,fedora,wordpress} to inject malware into packages.
 
@@
 
-Opam 2.0 is not released yet, but fixes several major concerns:
-downloading now validates the TLS certificates~\cite{opampr} using the system-wide trust anchors,
+Opam 2.0 is not released yet, but fixes major concerns:
+downloading validates the TLS certificates~\cite{opampr} using system-wide trust anchors,
 a wrap command is available for building and installing packages~\cite{wrappr}.
 Nevertheless, a breach of the central server enables an attacker to modify packages arbitrarily.
-Relying only on transport layer security for packages, where many implementations have a long history of security vulnerabilities (e.g. goto fail), is not optimal.
+Relying only on TLS for packages is suboptimal, because many TLS implementations suffer from a long history of security vulnerabilities (e.g. goto fail).
 
@@
 Using end-to-end signatures, neither the the server hosting the repository nor the server where the tarballs are downloaded from need to be trusted!
-Conex has been designed with opam in mind, but it can be used without opam to publish and update data in an authenticated way.
-At the moment, the main opam repository is a git repository hosted on GitHub.
-Conex, as presented here, uses GitHub identifiers, but it can use email addresses instead.
-Conex does not depend on GitHub, neither on git.
+Conex has been designed with opam in mind, but it can be used without opam to publish and update any data in an authenticated way.
 Conex is not limited to the main opam repository, but can be used with any public or private opam repository, or any other data repository.
@@
 The main goal of conex is to \emph{establish an authenticated way to distribute packages}.
-Conex is constrained by the current use of the opam repository:  most package authors release a package and submit a pull request (PR) to the main opam repository on GitHub.
-Afterwards, janitors (repository maintainers) check the PR (good package name, potentially missing dependencies or build-time only dependencies), automated builds are performed by TravisCI, and the automated bot Camelus~\cite{camelus} reports feedback and missing fields.
-Using conex, most of the burden has to be carried by the janitors.
+The main opam repository is a git repository hosted on GitHub.
+Conex does not depend on GitHub, or even on git.
+Conex is constrained by the current use of the main opam repository:  most package authors release a package and submit a pull request (PR).
+Janitors (repository maintainers) then check the PR, TravisCI builds it using several OCaml and opam versions, and the automated bot Camelus~\cite{camelus} reports opam inconsistencies.
 
 Conex is implemented in OCaml and licensed under the 2-clause BSD license (source~\cite{conex}).
-To provide authenticity from the start, conex should be installed at the same time as opam, thus it is constrained to use few dependencies, and will only depend on cmdliner and opam-lib.
+To provide authenticity from the beginning, conex is best installed together with opam.
 
-Earlier work on securing software update systems, such as the update framework (TUF)~\cite{tuf} and Diplomat~\cite{diplomat} focus on distributing directories securely, but make simultaneous updates of several packages by different authors hard, due to centralised files with authorisation and key information.
-Our earlier proposal for opam signing~\cite{opamsigning} adjusted ideas from TUF to the current opam workflow, but introduced superfluous structure.
+Earlier work on securing software update systems, such as the update framework (TUF)~\cite{tuf} and Diplomat~\cite{diplomat} focus on distributing directories securely, but due to centralised files simultaneous updates of packages are not free of conflicts.
+Our earlier proposal for opam signing~\cite{opamsigning} introduced unnecessary structure and complexity by adjusting TUF to the decentralised opam workflow.
 
-In section~\ref{sec:design} we will discuss the design of conex, followed by reasonable shortcuts in section~\ref{sec:shortcuts}.
-Afterwards in section~\ref{sec:implementation} we present the implementation and opam integration.
-We will discuss related work in section~\ref{sec:related} and conclude with future work in section~\ref{sec:conclusion}.
+In section~\ref{sec:design} we discuss the design of conex, followed by reasonable shortcuts in section~\ref{sec:shortcuts}.
+Afterwards in section~\ref{sec:implementation} we briefly present the implementation of cones and opam integration.
+Related work is mentioned in section~\ref{sec:related}, we conclude in section~\ref{sec:conclusion}.
 
 \section{Design} \label{sec:design}
-We first take a step back and do not consider all the details of opam for now.
-
 Conex does not claim to protect against undiscovered vulnerabilities in the packages, dependent tools, or conex itself.
@@
 \subsection{Threat model}
-There are many risks that users of software update systems face, ranging from interjecting traffic over weaknesses in TLS, weaknesses in the network infrastructure, compromising signing keys, stealing keys, to compromising the repository or signing infrastructure.
+There are many risks that users of software update systems face, ranging from injecting traffic over weaknesses in TLS, weaknesses in the network infrastructure, compromising signing keys, to compromising the repository or signing infrastructure.
 
-This leads us to consider a threat model where parts of the involved systems are compromised.  We assume that an attacker can:
+This leads us to consider a threat model where some involved systems are compromised.  We assume that an attacker can:
 \begin{itemize}
@@
 
-\begin{figure}
+\begin{figure}[h]
 \textbf{id}: restricted string\\
@@
 
-\begin{figure}
-  \centering
+\begin{figure}[h]
   \begin{minipage}{\hsize}
@@
     \end{minipage}
+    \hspace*{.5cm}
     \begin{minipage}{.4 \hsize}
@@
     \end{minipage}
+    \hspace*{.5cm}
     \begin{minipage}{.55 \hsize}
@@
 
-The releases file format, shown in Figure~\ref{fig:releasesformat}, contains the package name and the list of all releases of this package.
-It is signed by an authorised author (or a quorum of janitors).
+The releases file format (Figure~\ref{fig:releasesformat}) contains the package name and a list of all releases.
+It is signed by an authorised author or a quorum.
 This is crucial to avoid rollback attacks, where an attacker hinders a client from receiving package updates.
 
-The checksums file format, shown in Figure~\ref{fig:checksumformat}, contains the release name, and a list of file names which are part of the release (opam, url, descr, optionally patches, archive) and their sizes and digests.
-This list must contain all files recursively present in the current directory (excluding the checksum file itself).
+The checksums file format, shown in Figure~\ref{fig:checksumformat}, contains the release name, and a list of file names which are part of the release (opam, optionally patches, archive, ...), their sizes and digests.
+This list must contain all files recursively present in the current directory (excluding the checksum file).
 
-\begin{figure}
-  \centering
+\begin{figure}[h]
 counter: \textbf{Int64},\\
@@
 The set for the main opam repository is distributed with opam.
-The initial authenticated opam installation, including the initial set of janitors, is out of scope for conex (requires another package manager, OpenPGP signatures, ...).
+The opam installation, including the initial set of janitors, is out of scope for conex (e.g. by another package manager, OpenPGP signatures).
 
@@
 If the work overhead for janitors turns out too high, we can leverage the system in a way that \emph{neither authorisation nor public keys require a quorum}.
-This will open the system to name squatting attacks for identifiers.
-In a snapshot authorisations cannot be verified anymore, because there is no history to the authorisation.
-Update verification still applies: if we trust the timestamp notaries and use git, janitors don't have to sign public keys and authorisation files.
+This will open the system to name squatting attacks.
+Snapshot verification does not apply anymore, because there is no history of authorisations.
+Update verification can still be done, thus we could trust the timestamp notaries and use git.
 
@@
 Conex is work-in-progress~\cite{conex}.
-It does not depend on opam-lib, but instead will be functorised over the data layout (json, opam, s-expression).
-To further minimise the dependencies, the goal is to have a cryptographic module which uses the widely deployed \emph{openssl} utility for bootstrapping.
+It will be functorised over the data layout (json, opam, s-expression), and over a cryptography provider.
+One provider will be nocrypto, another the \emph{openssl} utility, enabling bootstrapping without opam.
+The timestamp notary and other automated tools will use the conex library.
 
-The timestamp notary and other automated tools will use the conex library as a basis.
-
-Conex is not yet integrated into opam, but some work on opam 2.0, such as compilers-as-packages, will make the integration smooth.
-The idea is to use conex as a separate application which proxies discover and download commands done by opam.
-
-There is still demand for more tooling support for janitors.
-In order to reach a quorum, individual janitors have to add the digest to their index file.
-Once the quorum is reached, the proposed PR can be merged.
-A queue of updates needing attention (a quorum) will be provided by the tooling.
+Conex is not yet integrated into opam, but work on opam 2.0, such as compilers-as-packages, are prerequisites for conex.
+There is still demand for decent tooling support for janitors, such as a queue of updates requiring signatures for a quorum.
 
 \section{Related work} \label{sec:related}
-Conex is based on the update framework~\cite{tuf}, which centralises all packages into a single \texttt{target.txt} file, and has more structure in keys.
-Instead of doing this, conex was designed with version control and distributed updates in mind.
+Conex is based on the update framework~\cite{tuf}, which centralises all packages into a single \texttt{target.txt} file.
+Instead of doing this, conex was designed with distributed updates in mind.
 
-A more recent contribution is Diplomat~\cite{diplomat}, but instead of automatically promoting unclaimed packages to claimed (which we can do using timestamp services), we believe in end-to-end signing of packages.
-Our security model relies on offline keys which are protected by authors.
+More recent is Diplomat~\cite{diplomat}, which automatically promoting unclaimed packages to claimed (which we can do using timestamp notary).
+Our security model instead uses end-to-end signatures of packages, and relies on authors to sign their packages.
 
-The Haskell signing proposal and implementation does not enforce authors to sign their packages, but instead the repository is signed.
-Thus, the repository server (or key server) needs to be trusted ultimately.
+The Haskell signing proposal does not enforce authors to sign their packages, but the repository is signed with an online key.
 
-Our original opam signing proposal~\cite{opamsigning} did not include a list of releases per package, opening up to attacks where the user does not receive a complete list of releases.
-Additionally, the original proposal introduced an unnecessary hierarchy of keys, and complicated the setup.
-It relied on the timestamping notary or to use git as update protocol with appropriate protection of the transport layer.
+Our original opam signing proposal~\cite{opamsigning} did not include a list of releases per package, opening up to various attacks.
+Additionally, our original proposal introduced an unnecessary hierarchical structure.
+It relied on the timestamping notary or using git as update protocol with appropriate protection of the transport layer.
 
@@
 We presented conex, which secures the software update system opam with minor modifications to the current workflow.
-Conex uses end-to-end signatures (from package authors to users) as trust model.
-It removes any trust into a central server.
-Known attacks, observed in practise or described in literaturre, are mitigated sufficiently by conex.
+Conex uses signatures from package authors to users, removing the demand to trust a central server.
+Conex mitigates known practical attacks as well as theoretical ones described in literature.
 
 Future work includes finishing the implementation of conex, integrating it into opam, and develop automated services (timestamp notary, build bot, PR bot).
+
+%\TODO{Acknowledge Joe, Raphael}
 

### 2016-06-09 paper.tex
Just a few minor remarks

@@
 downloading validates the TLS certificates~\cite{opampr} using system-wide trust anchors,
-a wrap command is available for building and installing packages~\cite{wrappr}.
+a wrap command is available for building and installing packages~\cite{wrappr}, optionally enabling sandboxing.
 Nevertheless, a breach of the central server enables an attacker to modify packages arbitrarily.
@@
 
+% LG: What about injecting packages (dependencies) into the user system ? I have
+% \texttt{provides:} in mind, that may enable this without changing the contents
+% of a package
+
 We cannot detect compromises of (offline!) keys, but this requires janitors, authors, and other people to carefully read through updates.
@@
 Each author updates their own packages, which should be free of conflicts due to separate resources for each package.
+% unclear without VC in mind, which is'nt introduced yet: precise ``conflicts between concurrent repository updates'' ?
 If a conflict is detected, it is up to the janitors to decide which update should go through, and which should be resubmitted on top of the other.
@@
 
-The \texttt{keys} directory contains one file for each author, named after the key identifier.
+The \texttt{keys} directory contains one file for each author and janitor, named after the key identifier.
 The \texttt{sigs} directory contains one file for each janitor, named after the janitor key identifier.
@@
 This would otherwise allow a detected attacker who compromised a key to deliver metadata signed with this compromised key.
+% LG: I am not sure I really get your meaning here
 

### 2016-06-09 paper.tex
sg + more concise

@@
 In section~\ref{sec:design} we discuss the design of conex, followed by reasonable shortcuts in section~\ref{sec:shortcuts}.
-Afterwards in section~\ref{sec:implementation} we briefly present the implementation of cones and opam integration.
+Afterwards in section~\ref{sec:implementation} we briefly present the implementation of conex and opam integration.
 Related work is mentioned in section~\ref{sec:related}, we conclude in section~\ref{sec:conclusion}.
@@
 
-Each author updates their own packages, which should be free of conflicts due to separate resources for each package.
+Each author updates their own packages, which should be free of conflicts between concurrent repository updates, due to separate resources for each package.
 % unclear without VC in mind, which is'nt introduced yet: precise ``conflicts between concurrent repository updates'' ?
@@
 Each file contains a \texttt{signed} structure: \texttt{(signed: }\textbf{resource}\texttt{, sigs: }\textbf{(id, signature) list}\texttt{)}.
+While signatures from authorised authors are done directly on the checksum/release/public key files, each janitor signs only their per-janitor index file.
+This reduces the amount of signature verifications which are done by the client.
 The different resources are explained in more detail below.
@@
 \textbf{role}: \texttt{Janitor} $|$ \texttt{Author} $|$ \texttt{Other} of \textbf{id}\\
-\textbf{service}: \texttt{EMail} of \textbf{name} $|$ \texttt{GitHub} of \textbf{name} $|$ \texttt{Other} of \textbf{id} * \textbf{name}\\
+\textbf{service}: \texttt{Email} of \textbf{name} $|$ \texttt{GitHub} of \textbf{name} $|$ \texttt{Other} of \textbf{id} * \textbf{name}\\
 \textbf{digest}: SHA256 digest of file\\
@@
 The data types for version 0 of conex are presented in Figure~\ref{fig:datatypes0}.
-A \texttt{restricted string} is a case insensitive 7 bit ASCII string without control characters or special characters.
+A \texttt{restricted string} is a case insensitive 7 bit ASCII string without control or white space characters.
 Depending on usage, more characters are forbidden (e.g. opam package names may not contain a '.').
@@
 The janitor index file format, shown in Figure~\ref{fig:janindexformat}, contains the janitor identifier, and all resources the janitor vouches for.
-
-Instead of removing author ids and packages, we put empty placeholders signed by janitors into the repository.
+It may contain any checksums, even of resources not yet in the repository (but proposed in some pull request).
+This avoids merge conflicts if a janitor wants to vouch for multiple new authors and packages at the same time.
+%\TODO{to avoid conflicts: it may contain checksums of resources not yet in the repository (in this way, multiple PRs may exist, which are merged once quorum is reached, but individual janitors can put the checksums into their index upfront!)}
 
@@
 The authorisation file itself has to be signed by a quorum of janitors.
-Public keys have to be signed by a quorum of janitors as well.
+Each public key has to be signed by its private key and a quorum of janitors.
 
@@
 \subsection{Verification}
-Both a snapshot of the repository can be verified using a set of initial trust anchors (public keys of janitors), and an update (patch) to an existing verified repository which leads to a verified repository with the patch applied.
+Since these verification steps are done by automated system (see Section~\ref{sec:shortcuts}) as well, verification errors are not expected at the client and treat as fatal.
+The user is informed about the potential security breach, the set of patches are not applied locally, and the user is asked to inform the janitors via email.
+Certainly, the user can try again to update their repository.
 
-\paragraph{Snapshot verification}  Verification of a snapshot is straightforward: take the set of initial janitors as trusted, then iterate through all packages, verify the authorisation file (properly signed by a quorum of janitors), all releases files, and verify all checksum files.
+A client can verify a repository in two ways:
+\begin{itemize}
+  \item a \emph{complete snapshot} of the repository (fresh clone)
+  \item an already verified repository and a list of \emph{updates} which should be applied
+\end{itemize}
+
+We will briefly outline the algorithm for both subsequently.
+
+\paragraph{Snapshot verification}  Verification of a snapshot is straightforward: take the set of initial janitor keys as trusted, verify all janitor index files, iterate over all packages, verify the authorisation file (properly signed by a quorum of janitors), verify all releases files, and verify all checksum files.
+The client also needs to ensure that there are no subdirectories or files which are not listed in the releases or checksum files.
 The public keys are verified upon demand: whenever a signature is encountered, the key of the signing id is verified.
 
-\paragraph{Update verification}  Consider that the repository in the old state is verified, use it to verify the update (in form of a patch) in the following way:
-\begin{itemize}
-  \item key modification -- key is sufficiently big, unique id matching filename, counter is increased, signed by quorum, no signature with old key
-  \item key insertion -- key is sufficiently big, unique id matching filename, counter is 0, signed by quorum
-  \item key deletion -- counter increasing, no signatures with this key, empty string is used to denote no key, signed by quorum
-  \item role modification -- counter increasing, no signature with old role, signed by quorum
-  \item authorisation insertion -- counter is 0, id matches directory, signed by quorum, authorised key ids are present
-  \item authorisation modification -- counter increasing, all checksum are signed with an authorised key, signed by quorum
-  \item complete package removal -- empty authorised, empty releases, signed by quorum
-  \item package insertion -- valid authorisation, releases, and checksum, both counter 0, signed by author (authorisation by quorum)
-  \item package update -- valid checksum, counter increasing, signed by author
-  \item new package version release -- index counter increasing and new version added to list, valid checksum, signed by author
-  \item package version deletion -- counter increasing, removed from releases, signed by author
-\end{itemize}
+\paragraph{Update verification}  The repository is already verified.
+The client uses it to verify each update separately.
+Each modified resource must be signed, either by an authorised author or by a quorum of janitors.
+The counter must start at 0 (upon creation) and increment.
+The file name and identifier must match.
+A key rollover needs to introduce the new key (replacing the old), and sign again all resources which were signed with the old key.
+Removal of resources is done by replacing it with the empty string.
+A new package needs to have an authorisation file and a releases file, both properly signed.
+%% \begin{itemize}
+%%   \item janitor index modification
+%%   \item janitor index creation - counter 0, properly named, signed by janitor
+%%   \item janitor index deletion - counter increasing, properly named, signed by quorum
+%%   \item key modification -- key is sufficiently big, unique id matching filename, counter is increased, signed by quorum, no signature with old key
+%%   \item key insertion -- key is sufficiently big, unique id matching filename, counter is 0, signed by quorum
+%%   \item key deletion -- counter increasing, no signatures with this key, empty string is used to denote no key, signed by quorum
+%%   \item role modification -- counter increasing, no signature with old role, signed by quorum
+%%   \item authorisation insertion -- counter is 0, id matches directory, signed by quorum, authorised key ids are present
+%%   \item authorisation modification -- counter increasing, all checksum are signed with an authorised key, signed by quorum
+%%   \item complete package removal -- empty authorised, empty releases, signed by quorum
+%%   \item package insertion -- valid authorisation, releases, and checksum, both counter 0, signed by author (authorisation by quorum)
+%%   \item package update -- valid checksum, counter increasing, signed by author
+%%   \item new package version release -- index counter increasing and new version added to list, valid checksum, signed by author
+%%   \item package version deletion -- counter increasing, removed from releases, signed by author
+%% \end{itemize}
 
@@
 
-Package metadata, signatures, and public keys are distributed in the same repository to prevent attackers from hindering one communication link, e.g. the public key one.
-This would otherwise allow a detected attacker who compromised a key to deliver metadata signed with this compromised key.
-% LG: I am not sure I really get your meaning here
+If package metadata and public keys are distributed from different sources, an attacker who compromised a key could hinder a user from receiving the key revocation by preventing communication to the key server.
+This would allow the attacker to continue signing packages with the compromised key and having the user install arbitrary data.
+To avoid this, package metadata, signatures, and public keys are in the same repository.
+%Package metadata, signatures, and public keys are distributed in the same repository to prevent attackers from hindering one communication link, e.g. the public key one.
+%This would otherwise allow a detected attacker who compromised a key to deliver metadata signed with this compromised key.
+%% LG: I am not sure I really get your meaning here
+% HM: hopefully now it is better!?
 
@@
 Future work includes finishing the implementation of conex, integrating it into opam, and develop automated services (timestamp notary, build bot, PR bot).
+A first beta version of conex, including opam integration, is expected to be available for the OCaml workshop 2016.
 

### 2016-06-10 paper.tex
more comments from r; some minor adjustments to condense it a bit more

@@
 
-\authorinfo{Hannes Mehnert\thanks{Funded by EPSRC Programme Grant EP/K008528/1}}{University of Cambridge}{}
+\authorinfo{Hannes Mehnert}{University of Cambridge}{}
 \authorinfo{Louis Gesbert}{OCamlPro}{}
@@
   Using conex, neither the distribution server nor the transport layer need to be trusted.
-  Authors and users need to adapt their workflow only slightly (or use opam-publish), whereas janitors (repository maintainers) need to use more sophisticated tools.
+  Authors and users need to adapt their workflow only slightly (or use opam-publish), whereas janitors (repository maintainers) need more tooling.
 \end{abstract}
@@
 
-In section~\ref{sec:design} we discuss the design of conex, followed by reasonable shortcuts in section~\ref{sec:shortcuts}.
+In section~\ref{sec:design} we discuss the design of conex, followed by reasonable system optimisations in section~\ref{sec:shortcuts}.
 Afterwards in section~\ref{sec:implementation} we briefly present the implementation of conex and opam integration.
@@
 \begin{itemize}
-  \item An \emph{author} who develops packages and releases them to the repository.
-  \item A \emph{janitor} who keeps the repository tidy and in a working state (fix up reverse dependencies etc.).  Janitors also maintain the package name and author name resources.
+  \item An \emph{author} who develops packages and releases them to the repository.  Each author signs their own packages.
+  \item A \emph{janitor} who keeps the repository tidy and in a working state (fix up reverse dependencies etc.).
+Janitors also maintain the package name and author name resources by signing them.
+Authorisations, that an author is responsible for a package, are signed by janitors as well.
+Instead of trusting a single janitor to sign these resources, a quorum of signatures of janitors is required to avoid this concentration of power.
 \end{itemize}
 
-An author signs their own packages.
-A janitor introduces authors, authorise authors for their packages, and also adjust packages due to updated dependencies (such as the version constraints, or minor patches to build on another platform).
 It is likely that a janitor acts as an author for their packages, there is no need for a janitor to have multiple private keys.
-In section~\ref{sec:shortcuts} we discuss ways to make conex easier to use, leveraging the security level.
-
-If a single janitor would be trusted to tidy up packages, compromising only one janitor key would be sufficient to undermine the security of conex.
-We avoid this by requiring a quorum of janitors to sign hotfixes.
 
@@
 The distinction between an author and a janitor is done via the role of the author information.
-For each janitor the repository contains a resource with all resources and checksums, signed by the janitor (to avoid way too many public key operations during sign and verify).
+The repository contains one index per janitor with all resources and checksums the janitor vouches for.
 
@@
 
-There are two universes of names which we need to protect against squatting attacks (including typographic ones~\cite{typo}): package names and key identifiers.
-Janitors need to faithfully verify new package and public key submissions.
+There are two universes of names which need to be protected against starvation attacks (and typographic ones~\cite{typo}): package names and key identifiers.
+Janitors need to carefully check new package and public key submissions.
 
@@
 \subsection{Repository layout}
-While conex is not bound to a file system, we'll reuse their terminology.
+Conex is not bound to the file system, but we use their terminology.
 The root directory has three subdirectories, \texttt{keys}, \texttt{sigs} and \texttt{data}.
@@
 \subsection{Verification}
-Since these verification steps are done by automated system (see Section~\ref{sec:shortcuts}) as well, verification errors are not expected at the client and treat as fatal.
+Since these verification steps are done by automated systems (see section~\ref{sec:shortcuts}) as well, verification errors are not expected at the client and treated as fatal.
 The user is informed about the potential security breach, the set of patches are not applied locally, and the user is asked to inform the janitors via email.
@@
 Otherwise, an attacker could replay e.g. an insecure package build recipe.
-Also, each resource includes the name, otherwise the signature of package A would be valid for package B.
+Also, each resource includes a name, otherwise the signature of package A would be valid for package B.
 Each signature contains the resource kind to avoid carrying over a signature from one resource to another.
@@
 
-\section{Shortcuts} \label{sec:shortcuts}
+\section{Optimisations} \label{sec:shortcuts}
 %\TODO{see also http://bentnib.org/posts/2016-04-12-authenticated-data-structures-as-a-library.html}
@@
 To mitigate mix-and-match attacks, a \emph{timestamp notary}, which updates the repository regularly, verifies all updates, and attaches its digital signature to the latest patch (via an git annotated tag over the git commit).
-The public key of the timestamp service is inside the repository and signed by multiple janitors.
-Clients can, instead of verifying packages and updates independently, trust the timestamp service (or multiple timestamp services).
+The public key of the timestamp notary is inside the repository and signed by multiple janitors.
+Clients can, instead of verifying packages and updates independently, trust the timestamp notary (or multiple timestamp notaries).
 This will reduce computational overhead (signature verification are expensive operations on big numbers), and might be worth for embedded devices.
-If there are multiple timestamp notaries run by different organisations, and its code is well audited, it might be reasonable to use them as default.
+If there are multiple instances of a timestamp notary run by different organisations, and its code is well audited, it might be reasonable to use them as default.
 While there is not yet an implementation of a timestamp notary, it will likely be implemented as a MirageOS~\cite{mirage,nqsb-tls} application, thus the trusted code base is small.
@@
 Conex is not yet integrated into opam, but work on opam 2.0, such as compilers-as-packages, are prerequisites for conex.
-There is still demand for decent tooling support for janitors, such as a queue of updates requiring signatures for a quorum.
+The opam publish utility will support authors: creating private keys, creating checksum, authorisation, and releases files, and signing these.
+There is a demand for tooling for janitors, such as a queue of updates requiring signatures for a quorum.
 
@@
 
-Future work includes finishing the implementation of conex, integrating it into opam, and develop automated services (timestamp notary, build bot, PR bot).
+Future work includes finishing the implementation of conex, integrating it into opam, and developing automated services (timestamp notary, build bot, PR bot).
 A first beta version of conex, including opam integration, is expected to be available for the OCaml workshop 2016.
 
-%\TODO{Acknowledge Joe, Raphael}
+\paragraph{Acknowledgements} Parts of this work were supported by EPSRC Programme Grant EP/K008528/1 (REMS - Rigorous Engineering of Mainstream Systems).
+We thank especially Joe and Rapha\"el for their feedback on earlier drafts.
 

### 2016-09-20 paper.tex
some updates to reflect the current code

@@
 Thus, a person-in-the-middle can inject arbitrary metadata (additional patches, modified packages, modified checksums, ...) during the discovery step.
-A package is downloaded using the URL specified in the metadata of the package (usually https, again without certificate validation), and checks the recorded MD5 digest (which is known to be weak~\cite{rfc6151}) against the downloaded archive.
+A package is downloaded using the URL specified in the metadata of the package (usually a secure channel using https, again without certificate validation), and checks the recorded MD5 digest (which is known to be weak~\cite{rfc6151}) against the downloaded archive.
 The build step is not contained into a temporary chroot environment, but may modify arbitrary files on disk.
-The recipe for building is part of the metadata, and may also include invoking any shell commands (such as downloading more things).
+The recipe for building is part of the metadata, and may also include invoking any shell commands (such as downloading more files and executing those).
 The install step may also modify arbitrary files.
@@
 a wrap command is available for building and installing packages~\cite{wrappr}, optionally enabling sandboxing.
-Nevertheless, a breach of the central server enables an attacker to modify packages arbitrarily.
-Relying only on TLS for packages is suboptimal, because many TLS implementations suffer from a long history of security vulnerabilities (e.g. goto fail).
+Nevertheless, a breach of the central repository server enables an attacker to modify packages arbitrarily.
+Relying only on TLS for packages is suboptimal, because many TLS implementations suffer from a long history of security vulnerabilities (e.g. goto fail, Heartbleed).
 
-In this paper we present \emph{conex}, a system which uses end-to-end digital signatures - the original author signs their package and build instructions, the client verifies this signature - to re-establish trust into packages installed by opam.
+In this paper we present \emph{conex}, a system which establishes end-to-end digital signatures - the original author signs their package release and build instructions, the client verifies this signature - to re-establish trust into packages installed by opam.
 Using end-to-end signatures, neither the the server hosting the repository nor the server where the tarballs are downloaded from need to be trusted!
@@
 The main goal of conex is to \emph{establish an authenticated way to distribute packages}.
-The main opam repository is a git repository hosted on GitHub.
+
+The main opam repository is currently a git repository hosted on GitHub.
 Conex does not depend on GitHub, or even on git.
@@
 
-We cannot detect compromises of (offline!) keys, but this requires janitors, authors, and other people to carefully read through updates.
+Conex cannot detect compromises of (offline!) keys.  But if janitors, authors, and other people carefully read through updates of packages, surprising releases will be detected.
 We assume that authors and janitors protect their private keys on their computers in a reasonable way, but we also have mechanisms in place for key revocation.
@@
 
-It is likely that a janitor acts as an author for their packages, there is no need for a janitor to have multiple private keys.
+A janitor may also act as an author, in case they develop and maintain packages on their own.  There is no need for a janitor to have multiple private keys.
 
@@
 We distinguish various resources which are all hosted in the same repository:
-package metadata (checksums, build recipe, dependencies), author information (public key, id, role), list of releases (of a single package), and authorisation (relation between package name and author).
-The distinction between an author and a janitor is done via the role of the author information.
-The repository contains one index per janitor with all resources and checksums the janitor vouches for.
+author information (public key, id), teams (members), package metadata (build recipe, dependencies, url to tarball), list of releases (of a single package), and authorisation (relation between package name and author).
+The repository contains one index per author with all resources and checksums the author vouches for.
+A janitor is a member of the team ``janitors''.
 
-Each resource is only valid if it is signed by either one of the authorised key identifiers, or a quorum of janitors.
+Each resource is only valid if it is signed by either one of the authorised key identifiers (in case of relesaes and package metadata checksums), or a quorum of janitors.
 Each resource contains a monotonic counter to prevent rollback attacks.
@@
 If a conflict is detected, it is up to the janitors to decide which update should go through, and which should be resubmitted on top of the other.
+The main source of conflict is likely team membership addition and deletion.
 
@@
 Conex is not bound to the file system, but we use their terminology.
-The root directory has three subdirectories, \texttt{keys}, \texttt{sigs} and \texttt{data}.
+The root directory has three subdirectories, \texttt{keys}, \texttt{index} and \texttt{packages}.
 
-The \texttt{keys} directory contains one file for each author and janitor, named after the key identifier.
-The \texttt{sigs} directory contains one file for each janitor, named after the janitor key identifier.
-The \texttt{data} directory consists of subdirectories, which each must contain a file \texttt{authorisation} and a file \texttt{releases}, plus possibly any subdirectories.
+The \texttt{keys} directory contains one file for each author and each team, named after their identifier.
+The \texttt{index} directory contains one index file for each author, named after the key identifier.  This contains a list of resources the author vouches for, and a digital signature.
+The \texttt{packages} directory consists of subdirectories, which each must contain a file \texttt{authorisation} and a file \texttt{releases}, plus possibly any subdirectories.
 Each subdirectory must be listed in the \texttt{releases} file, and must contain a \texttt{checksum} file besides any other data files and directories (all listed in the checksum file).
@@
 Any key-value representation which can be non-ambiguously normalised into a string, can be used (such as the opam format, json, s-expressions).
-Each file contains a \texttt{signed} structure: \texttt{(signed: }\textbf{resource}\texttt{, sigs: }\textbf{(id, signature) list}\texttt{)}.
-While signatures from authorised authors are done directly on the checksum/release/public key files, each janitor signs only their per-janitor index file.
-This reduces the amount of signature verifications which are done by the client.
+Each index file contains a \texttt{signed} structure: \texttt{(signed: }\textbf{resource}\texttt{, sigs: }\textbf{(id, signature) list}\texttt{)}.
+The amount of public key operations for verification is at most the number of authors of the repository.
 The different resources are explained in more detail below.
-The signature is computed over the normalised string representation of the resource concatenated with the public key id and the resource kind, separated by a space character.
+The signature is computed over the normalised string representation of the resource concatenated with the public key id, separated by a space character.
 
@@
 \textbf{public key}: RSA public key ($\geq$ 2048 bit, PKCS1, PEM encoded)\\
-\textbf{role}: \texttt{Janitor} $|$ \texttt{Author} $|$ \texttt{Other} of \textbf{id}\\
 \textbf{service}: \texttt{Email} of \textbf{name} $|$ \texttt{GitHub} of \textbf{name} $|$ \texttt{Other} of \textbf{id} * \textbf{name}\\
@@
 Depending on usage, more characters are forbidden (e.g. opam package names may not contain a '.').
-Distinction of id and name is only for clarity (ids are the public key ids, where names are the package names), they form two disjoint sets of identifiers.
-Upon insertion of new identifiers, checks for non-collision and validity need to be done.
+Distinction of id and name is only for clarity (ids are public key identifiers, names are package names), they form two disjoint sets of identifiers.
+Upon insertion of new identifiers, checks for non-collision and validity are done.
 
@@
 accounts: \textbf{service list},\\
-role: \textbf{role}
  \caption{\label{fig:keyformat} Public key}
@@
     \hspace*{.5cm}
+    \begin{minipage}{.4 \hsize}
+counter: \textbf{Int64},\\
+version: \textbf{Int64},\\
+name: \textbf{id},\\
+members: \textbf{id list}
+ \caption{\label{fig:teamformat} Team}
+    \end{minipage}
+  \end{minipage}
+\end{figure}
+
+The file format for a public key, presented in Figure~\ref{fig:keyformat}, contains the PEM encoded public key, the identifier (used for signatures and authorisations in opam), and accounts in other systems.
+
+A level of indirection is a team, shown in Figure~\ref{fig:teamformat}, which contains a team name and its members.  A package can be owned by public keys and teams.
+
+\begin{figure}[h]
+  \begin{minipage}{\hsize}
     \begin{minipage}{.4 \hsize}
@@
     \end{minipage}
-  \end{minipage}
-\end{figure}
-
-The file format for a public key, presented in Figure~\ref{fig:keyformat}, contains the PEM encoded public key, the identifier (used for signatures and authorisations in opam), accounts in other systems, and its role.
-
-The authorisations file format, shown in Figure~\ref{fig:authformat}, contains the package name and a list of authorised key identifiers for this package.
-
-\begin{figure}[h]
-  \begin{minipage}{\hsize}
-    \begin{minipage}{.3 \hsize}
+    \hspace*{.5cm}
+    \begin{minipage}{.4 \hsize}
 counter: \textbf{Int64},\\
@@
     \end{minipage}
-    \hspace*{.5cm}
-    \begin{minipage}{.55 \hsize}
+  \end{minipage}
+\end{figure}
+
+The authorisations file format, shown in Figure~\ref{fig:authformat}, contains the package name and a list of authorised identifiers, either refering to public keys or teams.
+
+Each releases file (Figure~\ref{fig:releasesformat}) contains the package name and a list of all releases.
+The purpose is to avoid rollback attacks on an individual package.
+It is signed by an authorised author or a quorum of janitors.
+
+\begin{figure}[h]
+  \begin{minipage}{\hsize}
+    \begin{minipage}{.4 \hsize}
 counter: \textbf{Int64},\\
@@
 name: \textbf{name},\\
-files: \textbf{(name, byte size, digest) list}
+files: \textbf{(name, size, digest) list}
  \caption{\label{fig:checksumformat} Checksum}
+    \end{minipage}
+    \hspace*{.5cm}
+    \begin{minipage}{.4 \hsize}
+counter: \textbf{Int64},\\
+version: \textbf{Int64},\\
+identifier: \textbf{id},\\
+resources: \textbf{(name, kind, digest) list}
+ \caption{\label{fig:janindexformat} Index}
     \end{minipage}
@@
 
-The releases file format (Figure~\ref{fig:releasesformat}) contains the package name and a list of all releases.
-It is signed by an authorised author or a quorum.
-This is crucial to avoid rollback attacks, where an attacker hinders a client from receiving package updates.
-
-The checksums file format, shown in Figure~\ref{fig:checksumformat}, contains the release name, and a list of file names which are part of the release (opam, optionally patches, archive, ...), their sizes and digests.
+The checksums file format, shown in Figure~\ref{fig:checksumformat}, contains the release name, and a list of file names which are part of the release (opam, optionally patches, ...), their sizes and digests.
 This list must contain all files recursively present in the current directory (excluding the checksum file).
 
-\begin{figure}[h]
-counter: \textbf{Int64},\\
-version: \textbf{Int64},\\
-identifier: \textbf{id},\\
-resources: \textbf{(name, kind, digest) list}
- \caption{\label{fig:janindexformat} Janitor index}
-\end{figure}
-
-The janitor index file format, shown in Figure~\ref{fig:janindexformat}, contains the janitor identifier, and all resources the janitor vouches for.
+The index file format, shown in Figure~\ref{fig:janindexformat}, contains the identifier, and all resources the author vouches for.
 It may contain any checksums, even of resources not yet in the repository (but proposed in some pull request).
-This avoids merge conflicts if a janitor wants to vouch for multiple new authors and packages at the same time.
+This avoids merge conflicts if an author wants to vouch for multiple versions of the same resources at the same time.
 %\TODO{to avoid conflicts: it may contain checksums of resources not yet in the repository (in this way, multiple PRs may exist, which are merged once quorum is reached, but individual janitors can put the checksums into their index upfront!)}
@@
   \item a \emph{complete snapshot} of the repository (fresh clone)
-  \item an already verified repository and a list of \emph{updates} which should be applied
+  \item an already verified repository and a list of \emph{updates} (in the form of patches) which should be applied
 \end{itemize}
@@
 
-\paragraph{Snapshot verification}  Verification of a snapshot is straightforward: take the set of initial janitor keys as trusted, verify all janitor index files, iterate over all packages, verify the authorisation file (properly signed by a quorum of janitors), verify all releases files, and verify all checksum files.
+\paragraph{Snapshot verification}  Verification of a snapshot is straightforward: take the set of initial janitor keys as trusted, verify all index files from this set, verify the janitors team, iterate over all packages, verify the authorisation file (properly signed by a quorum of janitors), verify all releases files, and verify all checksum files.
 The client also needs to ensure that there are no subdirectories or files which are not listed in the releases or checksum files.
@@
 
-\paragraph{Update verification}  The repository is already verified.
+\paragraph{Update verification}  The repository on disk is already verified.
 The client uses it to verify each update separately.
@@
 Conex is work-in-progress~\cite{conex}.
-It will be functorised over the data layout (json, opam, s-expression), and over a cryptography provider.
+It will be abstracted over the data layout (json, opam, s-expression), and over a cryptography provider.
 One provider will be nocrypto, another the \emph{openssl} utility, enabling bootstrapping without opam.
@@
 
+%Conex can also be viewed as a verified append-only log, similar to a blockchain.  Further research is needed to
+
 \section{Conclusion} \label{sec:conclusion}

### 2016-10-13 paper.tex
some updates

@@
 \begin{abstract}
-  Opam is a software update systems, responsible for discovering, downloading, building, and installing packages.
+  Opam is a software update system, responsible for discovering, downloading, building, and installing packages from community repositories.
   The data and metadata at each step should be authenticated to originate from a (transitively) trusted author.
   Opam does not include any mechanism to authenticate any step at the moment.
-  We propose \emph{conex}, which establishes end-to-end digital signatures from the package author to the user.
+  We propose \emph{conex}, which establishes digital signatures from the package author to the user.
   Using conex, neither the distribution server nor the transport layer need to be trusted.
-  Authors and users need to adapt their workflow only slightly (or use opam-publish), whereas janitors (repository maintainers) need more tooling.
+  Clients will install authenticated packages, and notice conex only if signatures cannot be verified.
+  Authors need to adapt their workflow slightly, an updated \emph{opam publish} will be developed.
+  Janitors (repository maintainers) need to adapt their workflow to emit cryptographic signatures to the repository (in addition to pressing merge PR buttons).
 \end{abstract}
@@
 \section{Introduction}
-Opam discovers, downloads, builds, and installs OCaml packages.
-To prevent injection of malware, all received data needs to be authenticated.
-A software update system relies on a central server which distributes a repository (package metadata: dependencies and build instructions) of packages.
-When a new package is released, it is inserted into the repository and distributed to all clients by the central server.
+Opam discovers, downloads, builds, and installs OCaml packages from community repositories.
+It is a software update system which relies on a central server which distributes a repository (package metadata: dependencies and build instructions) of packages.
+When a new package is released, it is inserted into the repository and distributed via the central server to all clients pulling it.
 The central server is an online service and a single point of failure.
@@
 
-In this paper we present \emph{conex}, a system which establishes end-to-end digital signatures - the original author signs their package release and build instructions, the client verifies this signature - to re-establish trust into packages installed by opam.
-Using end-to-end signatures, neither the the server hosting the repository nor the server where the tarballs are downloaded from need to be trusted!
+In this paper we present \emph{conex}, a system which establishes digital signatures - the original author signs their package release and build instructions, the client verifies this signature - to re-establish trust into packages installed by opam.
+Neither the the server hosting the repository nor the server where the tarballs are downloaded from need to be trusted!
 Conex has been designed with opam in mind, but it can be used without opam to publish and update any data in an authenticated way.
-Conex is not limited to the main opam repository, but can be used with any public or private opam repository, or any other data repository.
+Conex is not limited to the main opam repository, it can be used with any public or private opam repository.
 The only requirement is a second channel to distribute an initial set of trust anchors (e.g. an opam package in the main opam repository).
@@
 Conex is implemented in OCaml and licensed under the 2-clause BSD license (source~\cite{conex}).
-To provide authenticity from the beginning, conex is best installed together with opam.
+To provide authenticity for all other packages, conex is best installed together with opam.
+This leads to a constraint on conex: it may not have any dependencies apart from opam and the OCaml runtime.
+The installation of OCaml, opam, and conex need to be authenticated via other means (such as the OS package manager).
 
-Earlier work on securing software update systems, such as the update framework (TUF)~\cite{tuf} and Diplomat~\cite{diplomat} focus on distributing directories securely, but due to centralised files simultaneous updates of packages are not free of conflicts.
+Earlier work on securing software update systems, such as the update framework (TUF)~\cite{tuf} focusses on distributing directories securely, but due to centralised files simultaneous updates of packages are not free of conflicts.
+Diplomat~\cite{diplomat} was designed at a similar time as conex, and shares various goals as conex.
+%XXX: there need to be a more concrete comparison with diplomat somewhere
 Our earlier proposal for opam signing~\cite{opamsigning} introduced unnecessary structure and complexity by adjusting TUF to the decentralised opam workflow.
 
-In section~\ref{sec:design} we discuss the design of conex, followed by reasonable system optimisations in section~\ref{sec:shortcuts}.
+In section~\ref{sec:design} we discuss the design of conex, followed by a timestamping service in section~\ref{sec:shortcuts}.
 Afterwards in section~\ref{sec:implementation} we briefly present the implementation of conex and opam integration.
@@
 
-Conex cannot detect compromises of (offline!) keys.  But if janitors, authors, and other people carefully read through updates of packages, surprising releases will be detected.
+%(offline!) keys -- we need to be precise which keys are where..
+Conex cannot detect compromises of keys.  But if janitors, authors, and other people carefully read through updates of packages, releases not done by the original author will be detected.
 We assume that authors and janitors protect their private keys on their computers in a reasonable way, but we also have mechanisms in place for key revocation.
@@
   \item An \emph{author} who develops packages and releases them to the repository.  Each author signs their own packages.
-  \item A \emph{janitor} who keeps the repository tidy and in a working state (fix up reverse dependencies etc.).
+  \item A \emph{janitor} who keeps the repository in a working state (fix up reverse dependency constraints etc.).
 Janitors also maintain the package name and author name resources by signing them.
-Authorisations, that an author is responsible for a package, are signed by janitors as well.
 Instead of trusting a single janitor to sign these resources, a quorum of signatures of janitors is required to avoid this concentration of power.
@@
 A janitor may also act as an author, in case they develop and maintain packages on their own.  There is no need for a janitor to have multiple private keys.
+The set of janitors is a subset of the authors.
 
@@
 We distinguish various resources which are all hosted in the same repository:
-author information (public key, id), teams (members), package metadata (build recipe, dependencies, url to tarball), list of releases (of a single package), and authorisation (relation between package name and author).
-The repository contains one index per author with all resources and checksums the author vouches for.
+author information (public key, id, contact information), teams (members), package metadata (build recipe, dependencies, url to tarball), list of releases (of a single package), and authorisation (who is maintaining each package).
+For each author one signed index file is in the repository, containing all resources and digests the author vouches for.
 A janitor is a member of the team ``janitors''.
 
-Each resource is only valid if it is signed by either one of the authorised key identifiers (in case of relesaes and package metadata checksums), or a quorum of janitors.
-Each resource contains a monotonic counter to prevent rollback attacks.
+Each resource is only valid if it is vouched for either by an authorised key (in case of relesaes and package checksums), or a quorum of janitors.
+Each resource contains a counter, monotonically increasing between updates, to prevent rollback attacks.
+A quorum of janitors can reset the counter to recover from fast forward attacks.
 
@@
 If a conflict is detected, it is up to the janitors to decide which update should go through, and which should be resubmitted on top of the other.
-The main source of conflict is likely team membership addition and deletion.
+%Modifications to team membership can only be done sequentially 
+%The main source of conflict is likely team membership addition and deletion.
 
 \subsection{Repository layout}
-Conex is not bound to the file system, but we use their terminology.
+Conex is not bound to a file system, but we use their terminology.
 The root directory has three subdirectories, \texttt{keys}, \texttt{index} and \texttt{packages}.
@@
 The \texttt{keys} directory contains one file for each author and each team, named after their identifier.
-The \texttt{index} directory contains one index file for each author, named after the key identifier.  This contains a list of resources the author vouches for, and a digital signature.
-The \texttt{packages} directory consists of subdirectories, which each must contain a file \texttt{authorisation} and a file \texttt{releases}, plus possibly any subdirectories.
-Each subdirectory must be listed in the \texttt{releases} file, and must contain a \texttt{checksum} file besides any other data files and directories (all listed in the checksum file).
+The \texttt{index} directory contains one index file for each author, named after their key identifier.
+The \texttt{packages} directory consists of subdirectories, which each must contain an \texttt{authorisation} file and a \texttt{releases} file, plus possibly any subdirectories.
+Each subdirectory must be listed in the \texttt{releases} file, and must contain a \texttt{checksum} file besides any other data files and further subdirectories (all listed in the checksum file).
 
@@
 Any key-value representation which can be non-ambiguously normalised into a string, can be used (such as the opam format, json, s-expressions).
-Each index file contains a \texttt{signed} structure: \texttt{(signed: }\textbf{resource}\texttt{, sigs: }\textbf{(id, signature) list}\texttt{)}.
+Each index file contains a \texttt{signed} structure: \texttt{(signed: }\textbf{resource}\texttt{, sigs: }\textbf{(id, signature, timestamp) list}\texttt{)}.
 The amount of public key operations for verification is at most the number of authors of the repository.
 The different resources are explained in more detail below.
-The signature is computed over the normalised string representation of the resource concatenated with the public key id, separated by a space character.
+The signature is computed over the normalised string representation of the resource concatenated with the public key id, and the current timestamp, using a space character as separator.
 
@@
 \textbf{name}: restricted string\\
+\textbf{timestamp}: Seconds since UNIX epoch as \textbf{Int64}\\
 \textbf{signature}: Base64 encoded RSASSA-PSS with SHA256 (PKCS1)\\
@@
 \textbf{digest}: SHA256 digest of file\\
-\textbf{byte size}: size as \textbf{Int64} of file
+\textbf{byte size}: size as \textbf{Int64} of file\\
+\textbf{resource type}: PublicKey $|$ Team $|$ Checksum $|$ Releases $|$ Authorisation
 \caption{\label{fig:datatypes0}
@@
 
-Each file includes a data version (0 in this proposal) and a monotonic 64 bit counter (starting at 0).
+Each file includes a version (0 for this proposal) and a monotonic 64 bit counter (starting at 0).
 
@@
 identifier: \textbf{id},\\
-accounts: \textbf{service list},\\
- \caption{\label{fig:keyformat} Public key}
+accounts: \textbf{service list}
+ \caption{\label{fig:keyformat} Public key file}
     \end{minipage}
@@
 members: \textbf{id list}
- \caption{\label{fig:teamformat} Team}
+ \caption{\label{fig:teamformat} Team file}
     \end{minipage}
@@
 authorised: \textbf{id list}
- \caption{\label{fig:authformat} Authorisation}
+ \caption{\label{fig:authformat} Authorisation file}
     \end{minipage}
@@
 releases: \textbf{name list}
- \caption{\label{fig:releasesformat} Releases}
+ \caption{\label{fig:releasesformat} Releases file}
     \end{minipage}
@@
 files: \textbf{(name, size, digest) list}
- \caption{\label{fig:checksumformat} Checksum}
+ \caption{\label{fig:checksumformat} Checksum file}
     \end{minipage}
@@
 identifier: \textbf{id},\\
-resources: \textbf{(name, kind, digest) list}
- \caption{\label{fig:janindexformat} Index}
+resources: \textbf{(name, resource type, digest) list}
+ \caption{\label{fig:janindexformat} Index file}
     \end{minipage}
@@
 The file name and identifier must match.
-A key rollover needs to introduce the new key (replacing the old), and sign again all resources which were signed with the old key.
+A key rollover needs to introduce the new key (replacing the old), and a signature with the new key of the index file.
 Removal of resources is done by replacing it with the empty string.
-A new package needs to have an authorisation file and a releases file, both properly signed.
+A new package needs to have an authorisation file (signed by a quorum) and a releases file (signed by an authorised author).
 %% \begin{itemize}
@@
 Otherwise, an attacker could replay e.g. an insecure package build recipe.
+To prevent fast forward attacks, a quorum of janitors can reset counters.
 Also, each resource includes a name, otherwise the signature of package A would be valid for package B.
-Each signature contains the resource kind to avoid carrying over a signature from one resource to another.
+Each signature contains the resource type to avoid carrying over a signature from one resource to another.
 
 Instead of removing resources to revoke them, the already claimed names stay there.
-The public key, package, or janitor index resource is emptied and must be signed by a quorum of janitors.
+The public key or package is emptied and must be signed by a quorum of janitors.
 
-A person-in-the-middle attacker could implement a backdoor for a package, but would need to either modify the checksum, prepare a new release, embed the patch as metadata in the repository, or introduce a new package (which they could) and add a dependency.
+A person-in-the-middle attacker could implement a backdoor for a package, but would need to either modify the checksum, prepare a new release, embed the patch as metadata in the repository, or introduce a new package and add a dependency.
 All of these actions require either a signature from the original author or from a quorum of janitors.
@@
 They can be mitigated by relying on some append-only log, blockchain, or other monotonic updates, such as git without forced pushes.
-We will discuss a timestamp notary in the next section, mitigating both mix-and-match and freeze attacks.
+We will discuss a timestamping service in the next section, mitigating both mix-and-match and freeze attacks.
 
-\section{Optimisations} \label{sec:shortcuts}
+\section{Timestamping Service} \label{sec:shortcuts}
 %\TODO{see also http://bentnib.org/posts/2016-04-12-authenticated-data-structures-as-a-library.html}
@@
 
-To mitigate mix-and-match attacks, a \emph{timestamp notary}, which updates the repository regularly, verifies all updates, and attaches its digital signature to the latest patch (via an git annotated tag over the git commit).
-The public key of the timestamp notary is inside the repository and signed by multiple janitors.
-Clients can, instead of verifying packages and updates independently, trust the timestamp notary (or multiple timestamp notaries).
-This will reduce computational overhead (signature verification are expensive operations on big numbers), and might be worth for embedded devices.
-If there are multiple instances of a timestamp notary run by different organisations, and its code is well audited, it might be reasonable to use them as default.
-While there is not yet an implementation of a timestamp notary, it will likely be implemented as a MirageOS~\cite{mirage,nqsb-tls} application, thus the trusted code base is small.
+To mitigate mix-and-match attacks, a \emph{timestamp service}, which updates the repository regularly, verifies all updates, and attaches its digital signature to the latest patch.
+The public key of the timestamp service is part of the repository and signed by multiple janitors.
+If there are multiple instances of a timestamp service run by different organisations, and its code is well audited, it might be reasonable to only verify their signatures, and trust it to verify all author signatures.
+While there is not yet an implementation of a timestamp service, it will likely be implemented as a MirageOS~\cite{mirage,nqsb-tls} application, thus the trusted code base is small.
 
@@
 Initial set of janitors (can be a single element), the quorum (can be one) must be distributed for each other repository.
-Of course, an automated timestamp notary can be used for other repositories as well.
+Of course, an automated timestamp service can be used for other repositories as well.
 
 %\TODO{this is insecure, since an attacker can add themselves to the authorisation file and self-sign it!}
-If the work overhead for janitors turns out too high, we can leverage the system in a way that \emph{neither authorisation nor public keys require a quorum}.
+If the work overhead for janitors turns out too high, we can leverage the system in a way that \emph{neither authorisation nor public keys require a quorum of janitors}.
+Diplomat~\cite{diplomat} introduced online keys for this (their \texttt{new} role).
 This will open the system to name squatting attacks.
@@
 
-%that they contain the same information as the original ones, but updates can be verified (and again, we can outsource trust into the timestamping notary).
+%that they contain the same information as the original ones, but updates can be verified (and again, we can outsource trust into the timestamping service).
 
@@
 One provider will be nocrypto, another the \emph{openssl} utility, enabling bootstrapping without opam.
-The timestamp notary and other automated tools will use the conex library.
+The timestamp service and other automated tools will use the conex library.
 
@@
 
-More recent is Diplomat~\cite{diplomat}, which automatically promoting unclaimed packages to claimed (which we can do using timestamp notary).
-Our security model instead uses end-to-end signatures of packages, and relies on authors to sign their packages.
+More recent is Diplomat~\cite{diplomat}, which automatically promoting unclaimed packages to claimed (which we can do using timestamp service).
+Our security model instead uses digital signatures of packages, and relies on authors to sign their packages.
 
@@
 Additionally, our original proposal introduced an unnecessary hierarchical structure.
-It relied on the timestamping notary or using git as update protocol with appropriate protection of the transport layer.
+It relied on the timestamping service or using git as update protocol with appropriate protection of the transport layer.
 
@@
 
-Future work includes finishing the implementation of conex, integrating it into opam, and developing automated services (timestamp notary, build bot, PR bot).
+Future work includes finishing the implementation of conex, integrating it into opam, and developing automated services (timestamp service, build bot, PR bot).
 A first beta version of conex, including opam integration, is expected to be available for the OCaml workshop 2016.

### 2016-10-13 paper.tex
clarify here and there

@@
 Conex is not limited to the main opam repository, it can be used with any public or private opam repository.
-The only requirement is a second channel to distribute an initial set of trust anchors (e.g. an opam package in the main opam repository).
+The only requirement is a second channel to distribute an initial set of trust anchors.
 The main goal of conex is to \emph{establish an authenticated way to distribute packages}.
@@
 
+If a quorum of janitor keys is compromised, new trust anchors have to be generated and distributed.
+
 \section{Timestamping Service} \label{sec:shortcuts}
@@
 To mitigate mix-and-match attacks, a \emph{timestamp service}, which updates the repository regularly, verifies all updates, and attaches its digital signature to the latest patch.
-The public key of the timestamp service is part of the repository and signed by multiple janitors.
+The public key of the timestamp service can be part of the repository and signed by multiple janitors, or supplied via another channel (for private timestamp services).
 If there are multiple instances of a timestamp service run by different organisations, and its code is well audited, it might be reasonable to only verify their signatures, and trust it to verify all author signatures.
